### 2018-06-08 docs/paper/S1Table.Rmd
Made changes to manuscript body, just missing details on discussion

---
output: pdf_document
---

Update this table and fill with information from ERE's table

```{r setup, echo = FALSE}
library(magrittr)

read.csv(here::here("data", "gov_ind.csv")) %>% 
  knitr::kable(caption = "\\label{table:gov_ind} Indicators used to describe the social-ecological system", col.names = c("Indicator", "Isla Natividad", "Maria Elena", "Punta Herrero"), format = "latex") %>% 
  kableExtra::column_spec(1, width = "10em") %>% 
  kableExtra::column_spec(2, width = "8em") %>% 
  kableExtra::column_spec(3, width = "8em") %>% 
  kableExtra::column_spec(4, width = "8em") %>% 
  kableExtra::row_spec(1 , bold = T) %>% 
  kableExtra::row_spec(6 , bold = T) %>% 
  kableExtra::row_spec(14 , bold = T) %>% 
  kableExtra::row_spec(18 , bold = T)
```