### 2018-06-08 docs/paper/S5Table.Rmd

---
output: pdf_document
---

Update this table and fill with information from ERE's table

```{r, echo = F}
library(magrittr)
read.csv(here::here("data","gov_res.csv")) %>% 
  knitr::kable(caption = "\\label{table:gov_res}Analysis of the fishing cooperatives based on the Social-Ecological systems framework \\citep{mcginnis_2014}.",
               col.names = c("", "Indicator", "Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  kableExtra::column_spec(1, width = "10em") %>%
  kableExtra::column_spec(2, width = "10em") %>%
  kableExtra::row_spec(1 , bold = T) %>% 
  kableExtra::row_spec(7 , bold = T) %>% 
  kableExtra::row_spec(15 , bold = T) %>% 
  kableExtra::row_spec(19 , bold = T)
```