### 2018-06-08 docs/paper/S5Table.tex

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}

\begin{document}

Update this table and fill with information from ERE's table

\begin{verbatim}
## Warning in kableExtra::column_spec(., 1, width = "10em"): Please specify
## format in kable. kableExtra can customize either HTML or LaTeX outputs. See
## https://haozhu233.github.io/kableExtra/ for details.
\end{verbatim}

\begin{verbatim}
## Warning in kableExtra::column_spec(., 2, width = "10em"): Please specify
## format in kable. kableExtra can customize either HTML or LaTeX outputs. See
## https://haozhu233.github.io/kableExtra/ for details.
\end{verbatim}

\begin{verbatim}
## Warning in kableExtra::row_spec(., 1, bold = T): Please specify format in
## kable. kableExtra can customize either HTML or LaTeX outputs. See https://
## haozhu233.github.io/kableExtra/ for details.
\end{verbatim}

\begin{verbatim}
## Warning in kableExtra::row_spec(., 7, bold = T): Please specify format in
## kable. kableExtra can customize either HTML or LaTeX outputs. See https://
## haozhu233.github.io/kableExtra/ for details.
\end{verbatim}

\begin{verbatim}
## Warning in kableExtra::row_spec(., 15, bold = T): Please specify format in
## kable. kableExtra can customize either HTML or LaTeX outputs. See https://
## haozhu233.github.io/kableExtra/ for details.
\end{verbatim}

\begin{verbatim}
## Warning in kableExtra::row_spec(., 19, bold = T): Please specify format in
## kable. kableExtra can customize either HTML or LaTeX outputs. See https://
## haozhu233.github.io/kableExtra/ for details.
\end{verbatim}

\begin{longtable}[]{@{}lllll@{}}
\caption{\label{table:gov_res}Analysis of the fishing cooperatives based
on the Social-Ecological systems framework
\citep{mcginnis_2014}.}\tabularnewline
\toprule
& Indicator & Isla Natividad & Maria Elena & Punta
Herrero\tabularnewline
\midrule
\endfirsthead
\toprule
& Indicator & Isla Natividad & Maria Elena & Punta
Herrero\tabularnewline
\midrule
\endhead
Resource systems (RS) & & & &\tabularnewline
RS2 -- Clarity of system boundaries & TURF presence & High & High &
High\tabularnewline
RS3 -- Size of resource system & & & &\tabularnewline
RS5 -- Productivity of system & Type of ecosystem & High & High &
High\tabularnewline
RS7 -- Predictability of system dynamics & Intensity of environmental
disturbance & Low (ENSO) & High & High\tabularnewline
RS9 -- Location & Proximity to other communities/cities & Isolated & Not
Isolated & Not Isolated\tabularnewline
Governance systems (GS) & & & &\tabularnewline
GS1 -- Government organizations & Presence of fishing cooperatives & Yes
& Yes & Yes\tabularnewline
GS2 -- Nongovernment organizations & Involved actors & Yes & Yes &
Yes\tabularnewline
GS3 -- Network structure & Presence of an inter-cooperative structure &
Yes & No & No\tabularnewline
GS4 -- Property-rights systems & TURF presence & Yes & Yes &
Yes\tabularnewline
GS5 -- Operational-choice rules & Fishing Regulations / MPA enforcement
/ Enforcement technolofy & Yes & Yes & Yes\tabularnewline
GS6 -- Collective-choice rules & Cooperative regulations & Yes & Yes &
Yes\tabularnewline
GS7 -- Constitutional-choice rules & & & &\tabularnewline
Resource units (RU) & & & &\tabularnewline
RU1 -- Resource unit mobility & Targeted species home range & Low &
Medium & Medium\tabularnewline
RU2 -- Growth or replacement rate & Max age of targeted species & Low &
Medium & Medium\tabularnewline
RU4 -- Economic value & Price of targeted species & high & High &
high\tabularnewline
Actors (A) & & & &\tabularnewline
A1 -- Number of relevant actors & & 98 & &\tabularnewline
A2 -- Socioeconomic attributes & & & &\tabularnewline
A5 -- Leadership/entrepreneurship & Leadership & High & High &
High\tabularnewline
A6 -- Norms (trust-reciprocity)/social capital--- (Based on illegal
fishing) & Level of illegal fishing & High & High & Low\tabularnewline
A8 -- Importance of resource (dependence) & Presence of alternative
livelihoods & High & High & High\tabularnewline
\bottomrule
\end{longtable}

\end{document}