### 2018-09-18 docs/paper/Villasenor-Derbez.tex

\documentclass{frontiersSCNS}
\usepackage{url,hyperref,lineno,microtype,subcaption}
\usepackage[onehalfspacing]{setspace}
\usepackage{float}

\linenumbers

\usepackage[utf8]{inputenc}
\floatplacement{figure}{H}

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,2,*}\), Eréndira
Aceves-Bueno\(^{1}\), Álvin Suarez\(^{2}\), Stuart Fulton\(^{2}\),
Arturo Hernández-Velasco\(^{2}\), Jorge Torre\(^{2}\), Fiorenza
Micheli\(^{3}\)}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
of California, Santa Barbara, Santa Barbara, CA,
USA\newline \(^{2}\)Comunidad y Biodiversidad A.C., Guaymas, Sonora,
Mexico\newline \(^{3}\)Hopkins Marine Station and Center for Ocean
Solutions, Stanford University, Pacific Grove, CA, USA}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California,
Santa Barbara, Santa Barbara, CA, 93106}

\def\corrEmail{\href{mailto:jvillasenor@bren.ucsb.edu}{\nolinkurl{jvillasenor@bren.ucsb.edu}}}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}
\onecolumn
\firstpage{1}

\title[Community-based marine reserves]{Effectiveness of community-based marine reserves in small-scale
fisheries}

\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}

\maketitle

\begin{abstract}

Coastal marine ecosystems provide livelihoods for small-scale fishers
and coastal communities around the world. Artisanal fisheries face great
challenges since they are difficult to monitor, enforce, and manage.
Combining territorial user rights for fisheries (TURF) with no-take
marine reserves to create TURF-reserves is believed to improve the
performance of small-scale fisheries by buffering fisheries from
environmental variability and management errors, while ensuring that
fishers reap the benefits of conservation investments. In the last six
years, and following a 2012 regulation, 18 TURF-reserves have been
implemented in Mexico; their effectiveness has not been formally
evaluated. We combine causal inference techniques and a
social-ecological systems framework to provide a holistic evaluation of
community-based TURF reserves in three coastal communities in Mexico. We
find that while reserves have not yet achieved their stated goal of
increasing lobster densities, they continue to receive significant
support from the fishing communities. A lack of ecological and
socioeconomic effects likely results from a combination of factors.
First, the lobster fisheries are already well managed, and it is
unlikely that reserves might have a detectable effect. Second, some of
the reserves are not large enough to protect lobsters' home ranges.
Third, some of these reserves might be too young for the effects to
show. However, these reserves have shaped small-scale fishers' way of
thinking about marine conservation, which can provide a foundation for
establishing additional, larger marine reserves needed to effectively
conserve mobile species.

\medskip
\tiny
 \keyFont{ \section{Keywords:} TURF-reserves, Causal Inference, Social-Ecological Systems, Marine
Protected Areas, Marine Conservation, Small-Scale Fisheries}

\end{abstract}

\clearpage

\section{Introduction}\label{introduction}

Marine ecosystems around the world sustain significant impacts due to
overfishing and unsustainable fishing practices
\citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. In particular,
artisanal fisheries face great challenges since they tend to be hard to
monitor and enforce \citep{costello_2012}. Recent research shows that
combining Territorial Use Rights for Fisheries (TURFs) with no-take
marine reserves (MR) can greatly improve the performance of coastal
fisheries and the health of the local resouces
\citep{costello_2010-Ix,lester_2017}. Commonly known as TURF-Reserves,
these systems increase the benefits of spatial access rights allowing
the maintainance of healthy resources
\citep{afflerbach_2014-HP,lester_2017}. Although in theory these systems
are successful \citep{costello_2010-Ix}, there is little empirical
evidence of their effectiveness and the drivers of their success
\citep{afflerbach_2014-HP,lester_2017,smallhornwest_2018}.

The performance of these systems depends on how environmental and social
factors combine and interact. The science of marine reserves has largely
focused on understanding the ecological effects of these areas, which
include increased biomass, species richness, and densities of organisms
within the protected regions, climate change mitigation, and protection
from environmental variability
\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
Modelling studies show that fishery benefits of marine reserves depend
on initial stock status and the management under which the fishery
operates, as well as reserve size and the amount of larvae exported from
these \citep{hilborn_2006,krueck_2017-J1,deleo_2015}. Other research has
focused on the relationship between socioeconomic and governance
structures and reserve effectiveness
\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. However, to our
knowledge, no studies exist that evaluate TURF-reserves from both a
social and ecological perspective. This is especially important in
social-ecological coastal systems dominated by close interaction and
feedbacks between people and natural resources \citep{ostrom_2009-hg}.

TURF-reserves can be created as community-based marine reserves,
voluntarily established and enforced by local communities. This
bottom-up approach increases compliance and self-enforcement
\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
Community-based spatial closures occur in different contexts, like the
\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
\citep{bohnsack_2004,johannes_2002}. However, without legal recognition
no-take regulations are difficult to enforce and fishers rely on the
exclusive access granted by the TURF. In an effort to bridge this
normative gap, Civil Society Organizations (CSOs) have served as a link
between fishers and government, and set out to create a legal framework
that solve this governance issue. In Mexico, a new norm was created in
2014 allowing fishers to request the legal recognition of
community-based reserves as ``Fish Refuge'' (\emph{Zona de Refugio
Pesquero}; \citet{nom}). Fish refuges can be implemented as temporal or
partial reserves, which can protect one, some, or all resources within
their boundaries. Since 2012, 45 of Fish Refuges have been created along
the Pacific, Gulf of California, and Mexican Caribbean coastlines, with
18 of them implemented as TURF-reserves. However, their effectiveness
has not yet been formally evaluated and reported in the scientific
literature.

Here, we combine causal inference techniques and a social-ecological
systems framework to provide a holistic evaluation of community-based
marine reserves in three coastal communities in Mexico. These three case
studies span a range of ecological and social conditions representative
of different regions of Mexico. The objective of this work is twofold.
First, to provide a triple bottom line evaluation of the effectiveness
of community-based marine reserves that can inform similar processes in
other countries. Second, to evaluate the effectiveness of TURF-reserves
established as Fish Refuges in Mexico to identify opportunities where
improvement or adjustment might lead to increased effectiveness. We draw
from lessons learned in these three case studies and provide management
recommendations to maximize the effectiveness of community-based marine
reserves in small-scale fisheries in Mexico and in other regions around
the world that are using this tool to manage and rebuild their coastal
fisheries.

\section{Methods}\label{methods}

\subsection{Study area}\label{study-area}

We evaluate three TURF-reserves in Mexico (Fig \ref{fig:map}A). The
first one was created by the \emph{Buzos y Pescadores de la Baja
California} fishing cooperative, located in Isla Natividad in the Baja
Peninsula (Fig \ref{fig:map}B). The main fishery in the island is the
spiny lobster (\emph{Panulirus interruptus}), but other resources like
finfish, sea cucumber, read sea urchin, snail, and abalone are also an
important source of income. In 2006, the community decided to implement
two marine reserves within their fishing grounds to protect commercially
important invertebrate species; mainly lobster and abalone. While these
reserves obtained legal recognition only in 2018, they have been well
enforced since their implementation.

The other two TURF-reserves are located in Maria Elena and Punta
Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). In contrast with
Isla Nativdad, which hosts a well established fishing community, Maria
Elena is a fishing camp --visited intermittently during the fishing
season-- belonging to the Cozumel fishing cooperative (\emph{SCPP
Cozumel}); Punta Herrero is home to the \emph{SCPP José María Azcorra}
cooperative, and similar to Isla Natividad hosts a local community.
Their main fishery is the Caribbean spiny lobster (\emph{Panulirus
argus}), but they also target finfish in the off-season. Maria Elena and
Punta Herrero established eight marine reserves in 2012, and four marine
reserves in 2013, respectively. All these reserves have been legally
recognized as Fishing Refuges since their creation
\citep{dof_website_2012,dof_website_2013}.

These communities are representative of their region in terms of
ecology, socioeconomic, and governance aspects. Isla Natividad, for
example, is part of a greater group of fishing cooperatives belonging to
a Federation of Fishing Cooperatives. This group has been identified as
a cohesive group that often cooperates to better manage their resources
\citep{mccay_2017-1m,mccay_2014-CN,acevesbueno_2017}. Likewise, Maria
Elena and Punta Herrero are representative of fishing cooperatives in
the Mexican Caribbean, which are also part of a regional Federation.
Together, these three communities provide an accurate representation of
other fishing communities in each of their regions. While each region
has additional communities that have established community-based
TURF-reserves, available data would not allow us to perform the in-depth
analysis that we undertake. Yet, given the similarities among
communities and the socioeconomic and governance setting under which
they opperate, it is safe to cautiously generalize our results to other
communities in Mexico and other regions around the world.

\clearpage

\begin{figure}
\centering
\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{\label{fig:unnamed-chunk-1}\label{fig:map}Location of the three
coastal communities studied (A). Isla Natividad (B) is located off the
Baja California Peninsula, Maria Elena and Punta Herrero (C) are located
in the Yucatan Peninsula. Blue polygons represent the TURFs, and red
polygons the marine reserves.}
\end{figure}

\subsection{Data collection}\label{data-collection}

We use three main sources of information to evaluate these reserves
across the ecological, socioeconomic, and governance dimensions.
Ecological data come from the annual ecological monitoring of reserve
and control areas, carried out by members from each community and
personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
(\href{www.cobi.org.mx}{COBI}). Trained divers record richness and
abundances of fish and invertebrate species along replicate transects
(30x 2 m each) at depths 5-20 m in the reserves and control sites
\citep{fulton_2018,fulton_2019}. Size structures are also collected
during fish surveys. We define control sites as regions with habitat
characteristics similar to the corresponding reserves, and that
presumably had a similar probability of being selected as reserves
during the design phase. We focus our evaluation on sites where data are
available for reserve and control sites, before and after the
implementation of the reserve. This provides us with a
Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that
allows us to capture and control for temporal and spatial dynamics
\citep{depalma_2018,ferraro_2006-oW}. BACI designs and causal inference
techniques have proven effective to evaluate marine reserves, as they
allow us to causally attribute observed changes to the intervention
\citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites were surveyed
annually, and at least once before implementation of the reserves. Table
\ref{table:com_sum} shows a summary of the TURF-reserves included in
this study.

\clearpage

\begin{table}[H]

\caption{\label{tab:unnamed-chunk-2}\label{table:com_sum} Summary of analyzed TURF-reserves by community.}
\centering
\begin{tabular}[t]{l|r|r|l|r}
\hline
Community & TURF area ($\mathrm{km}^{-2}$) & Reserve area ($\mathrm{km}^{-2}$) & Percent as reserves & Year of implementation\\
\hline
Isla Natividad & 889.5 & 1.53 & 0.17\% & 2006\\
\hline
Maria Elena & 353.1 & 0.10 & 0.03\% & 2012\\
\hline
Punta Herrero & 299.7 & 0.43 & 0.14\% & 2013\\
\hline
\end{tabular}
\end{table}

Socioeconomic data come from landing receipts reported to the National
Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de
Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings
(Kg) and revenues (MXP) for cooperatives with and without marine
reserves. Cooperatives incorporated in this analysis belong to larger
regional-level Cooperative Federations, and are exposed to the same
markets and institutional frameworks, making them plausible controls
\citep{mccay_2017-1m,ayer_2018}. Landings and revenues were aggregated
at the cooperative-year level, and revenues were adjusted to represent
2014 values by the Consumer Price Index for Mexico \citep{oecd_2017-VV}
as:

\begin{equation}
I_t = RI_t\times\frac{CPI_t}{CPI_T}
\label{eqn:cpi}
\end{equation}

Where \(I_t\) represents the adjusted income for year \(t\) as the
product between the reported income for that year and the ratio between
the consumer price index in that year (\(CPI_t\)) to the most recent
year's consumer price index (\(CPI_T\)).

Data for the operationalization of the social-ecological system were
collected at the community--level from official documents used in the
creation and designation of the marine reserves
\citep{dof_website_2012,dof_website_2013,dof_website_2018} and based on
the authors' experience and knowledge of the communities. These include
information on the resource system, the resource units, actors, and the
governance system itself (Table \ref{tab:ses}).

\subsection{Data analysis}\label{data-analysis}

We evaluate the effect that marine reserves have had on four ecological
and two socioeconomic indicators (Table \ref{table:indicators}). Recall
that reserves were implemented to protect lobster and other benthic
invertebrates. However, we also use the available fish data to test for
associated co-benefits.

\begin{table}[H]

\caption{\label{tab:unnamed-chunk-3}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
\centering
\begin{tabular}[t]{l|l|l}
\hline
Category & Indicator & Units\\
\hline
Biological & Lobster density & org $\mathrm{m}^{-2}$\\
\hline
Biological & Invertebrate density & org $\mathrm{m}^{-2}$\\
\hline
Biological & Fish biomass & Kg $\mathrm{m}^{-2}$\\
\hline
Biological & Fish density & org $\mathrm{m}^{-2}$\\
\hline
Socioeconomic & Income from target species & M MXP\\
\hline
Socioeconomic & Landings from target species & Metric Tonnes\\
\hline
\end{tabular}
\end{table}

\clearpage

We use a difference-in-differences analysis to evaluate these
indicators. This approach allows us to estimate the effect that the
reserve had by comparing trends across time and treatments
\citep{moland_2013-VP,Villasenor-Derbez_2018}. The analysis of
ecological indicators is performed with a multiple linear regression of
the form:

\begin{equation}
I_{itj} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \sigma_jSpp_j + \epsilon
\label{eqn:reg_bio}
\end{equation}

Where year-fixed effects are represented by \(\gamma_{it} Year_t\), and
\(\beta Zone_i\) captures the difference between reserve (\(Zone = 1\))
and control (\(Zone = 0\)) sites. The interaction term
\(\lambda_{it} Year_t\times Zone_i\) represents the mean change in the
indicator inside the reserve, for year \(t\), with respect to the year
of implementation in the control site (See Table \ref{table:com_sum}).
When evaluating biomass and densities of the invertebrate or fish
communities, we include \(\sigma_j\) to control for species-fixed
effects.

Socioeconomic indicators are evaluated with a similar approach. Due to
data constrains, we only evaluate socioeconomic data for Isla Natividad
(2000 - 2014) and Maria Elena (2006 - 2013). Neighboring communities are
used as counterfactuals that allow us to control for unobserved
time-invariants. Each focal community (Isla Natividad and Maria Elena)
has three counterfactual communities.

\begin{equation}
I = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i + \sigma_jCom_j +\epsilon
\label{eqn:soc_reg}
\end{equation}

The model interpretation remains as for Eq \ref{eqn:reg_bio}, but in
this case the \(Treated\) dummy variable indicates if the community has
a reserve (\(Treated = 1\)) or not (\(Treated = 0\)) and \(\sigma_jCom\)
captures community-level fixed-effects. These regression models allow us
to establish a causal link between the implementation of marine reserves
and the observed trends by accounting for temporal and spatial dynamics
\citep{depalma_2018}. The effect of the reserve is captured by the
\(\lambda_t\) coefficient, and represents the difference observed
between the control site before the implementation of the reserve and
the treated sites at time \(t\) after controlling for other time and
space variations (\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively).
All model coefficients were estimated via ordinary least-squares and
heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All
analyses were performed in R 3.5.0 and R Studio 1.1.453 \citep{R_2018}.
Data and code are available on
\href{https://github.com/jcvdav/ReserveEffect}{github.com}.

We use the social-ecologycal system (SES) framework to evaluate each
community as a means of providing an explanation to the biological and
socioeconomic results. The use the SES framework standardizes our
analysis and allows us to communicate our results in a common language
across fields. We based our variable selection primarily on
\citet{leslie_2015-na,basurto_2013-oq}, who have previously analyzed
Mexican fishing cooperatives using this framework. Table \ref{tab:ses}
shows the selected variables, the definition and selected indicators,
and scoring system.

We followed \citet{leslie_2015-na} to operationalize the implementation
of the SES framework. We analyzed 18 variables, that fitted within the
first, second and third tires of the framework. Categorical variables
were scored as binary data (0 or 1). Numerical variables were given a
continuous score, and then standardized between 0 and 1. For example,
the score on number of actors was relative to the smallest values (21
fishers) which was given a score of one since theoretically smaller
groups facilitate coordination \citep{viana_2018}. Each component of the
SES framework had a cumulative value of 1 and was equally weighted to
calculate the final score.

\section{Results}\label{results}

The following sections present the effect that marine reserves had on
each of the biological and socioeconomic indicators for each coastal
community. Results are presented in terms of the difference through time
and across sites, relative to the control site on the year of
implementation (\emph{i.e.} effect size \(\lambda_t\)). We also provide
an overview of the governance settings of each community, and discuss
how these might be related to the effectiveness and performance of the
reserves.

\subsection{Biological effects}\label{biological-effects}

Indicators showed ambiguous responses through time for each reserve.
Figure \ref{fig:indicators}A shows positive effect sizes for lobster
densities in Isla Natividad and Punta Herrero during the first years,
but the effect is eroded through time. In the case of Maria Elena,
positive changes were observed in the third and forth year. These
effects are in the order of 0.2 extra organisms \(\mathrm{m}^{-2}\) for
Isla Natividad and Punta Herrero, and 0.01 organisms \(\mathrm{m}^{-2}\)
for Maria Elena, but are not significantly different from zero
(\(p > 0.05\)). Likewise, no changes were detected in fish biomass or
invertebrate and fish densities (\ref{fig:indicators}B-D), where effect
sizes oscillated around zero without clear trends. Full tables with
model coefficients are presented in the supplementary materials (S1
Table, S2 Table, S3 Table).

\subsection{Socioeconomic effects}\label{socioeconomic-effects}

Lobster landings and revenue were only available for Isla Natividad and
Maria Elena (Fig \ref{fig:lobsters}). For all years before
implementation, the effect sizes are close to zero, indicating that the
control and treatment sites have similar pre-treatment trends,
suggesting that these are plausible controls. However, effect sizes do
not change after the implementation of the reserve. Interestingly, the
negative effect observed for Isla Natividad on year 5 correspond to the
2011 hypoxia events. The only positive change observed in lobster
landings is for Isla Natividad in 2014 (\(p < 0.1\)). The three years of
post-implementation data for Maria Elena do not show a significant
effect of the reserve. Isla Natividad shows higher revenues after the
implementation of the reserve, as compared to the control communities.
However, these changes are not significant and are associated to
increased variation. Full tables with model coefficients are presented
in the supplementary materials (S4 Table, S5 Table.

\subsection{Governance}\label{governance}

Although we have little information on the social dimension of these
fisheries, we can use the social-ecological systems framework to analyze
the performance of each governance system (Table \ref{tab:ses}). Our
analysis shows that all communities analyzed share similarities in their
Governance system which is based on cooperatives (GS5.2.3.2), with
strong rules in use that include Operational rules (GS6.2),
Collective-choice rules (GS6.3), Constitutional rules (GS6.3), and even
Territorial use communal rights (GS6.1.4.3). However, we identified
important differences in terms of the actors, resource systems, and
resource units. Although all communities show a high level of leadership
(A5), the level of trust (A6.1) is lower in Punta Herrero. In general,
the presence and success of conservation initiatives depends on the
incentives of local communities to maintain a healthy status of the
resources they depend upon \citep{jupiter_2017}. The enabling conditions
for conservation seem to be strongly present in all communities. Due to
the clarity of access rights and isolation, the benefits of conservation
directly benefit the members of the fishing cooperative. These
conditions have favored the development of efficient community-based
enforcement systems.

\clearpage

\begin{figure}
\centering
\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{\label{fig:unnamed-chunk-4}\label{fig:indicators}Effect sizes for
marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME;
blue triangles), and Punta Herrero (PH; green squares) for lobster
densities (\emph{Panulirus spp}; A), fish biomass (B), invertebrate
densities (C), and fish densities (D). Plots are ordered by survey type
(left column: invertebrates; right column: fish). Points are jittered
hotizontally to avoid overplotting. Points indicate the effect size and
standard errors. Years have been centered to year of implementation.}
\end{figure}

\begin{figure}
\centering
\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{\label{fig:unnamed-chunk-7}\label{fig:lobsters}Effect sizes for
lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
circles) and Maria Elena (ME; blue triangles). Points indicate the
effect size and standard errors. Years have been centered to year of
implementation.}
\end{figure}

\section{Discussion}\label{discussion}

Our results indicate that these TURF-reserves have not increased lobster
densities. Additionally, no co-benefits were identified when using other
ecological indicators other than the previously reported buffering
effect that reserves can have to environmental variability in Isla
Natividad \citep{micheli_2012-EU}. The socioeconomic indicators
pertaining landings and revenues showed little to no change after
reserve implementation. The coastal ecosystems where these reserves are
located have been profoundly affected by climatic and oceanographic
extremes, including warming events, extreme storms and prolonged hypoxia
\citep{micheli_2012-EU}, (\textbf{Beas et al. in prep., Woodson et al in
review}). Despite the lack of evidence of the effectiveness of these
reserves, most of the communities show a positive perception about their
performance and continue to support their presence \citep{ayer_2018}.
Understanding the social-ecological context in which these communities
and their reserves operate might provide insights to this.

Some works evaluate marine reserves by performing inside-outside
\citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD} or
before-after comparisons \citep{betti_2017-lq}. The first approach does
not address temporal variability, and the second can not distinguish
between the temporal trends in a reserve and the entire system
\citep{depalma_2018}. Our approach to evaluate the temporal and spatial
changes provides a more robust measure of reserve effectiveness. For
example, we capture previously described patterns like the rapid
increase observed for changes in lobster densities for Isla Natividad on
the sixth year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A) occur a
year after the hypoxia events described by \citet{micheli_2012-EU},
which caused mass mortality of sedentary organisms such as abalone and
sea urchins, but not lobster and finfish. Yet, our empirical approach
assumes control sites are a plausible counterfactual for treated sites.
This implies that treated sites would have followed the same trend as
control sites, had the reserves not been implemented. Nonetheless,
temporal trends for each site don't show any significant increases,
supporting our findings of lack of change in the indicators used.

A first possible explanation for the lack of effectivenes may be the
young age of the reserves. Literature shows that age and enforcement are
important factors that influence reserve effectiveness
\citep{edgar_2014-UO,babcock_2010}. Isla Natividad has the oldest
reserves, and our SES analysis suggests that all communities have a
well-established community-based enforcement system. With these
characteristics, one would expect the reserves to be effective. Maria
Elena and Punta Herrero are relatively young reserves (\emph{i.e.}
\textless{} 5 years old) and effects may not yet be evident due to the
short duration of protection, relative to the life histories of the
protected species; other community-based marine reserves in tropical
ecosystems may take up to six years to show a spillover effect
\citep{dasilva_2015-zX}.

Another key condition for effectiveness is reserve size
\citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be
attributed to poor ecological coherence in reserve design (\emph{sensu}
\citet{rees_2018}. Previous research has shown that reserves in Isla
Natividad yield fishery benefits for the abalone fishery
\citep{rossetto_2015-V0}. Abalone are less mobile than lobsters, and
perhaps the reserves provide enough protection to these sedentary
invertebrates, but not lobsters. Design principles developed by
\citet{green_2017} for marine reserves in the Caribbean state that
reserves ``should be more than twice the size of the home range of
adults and juveniles'', and suggest that reserves seeking to protect
spiny lobsters should have at least 14 km across. Furthermore, may favor
implementation of reserves that pose low fishing costs due to their
small size or location. Our analysis of economic data supports this, as
neither landings nor revenues showed the expected short-term costs
associated to the first years of reserve implementation
\citep{ovando_2016-Wg}.

Even if reserves had appropriate sizes and were placed in optimal
locations, there are other plausible explanations for the observed
patterns. For instance, marine reserves are only likely to provide
fisheries benefits if initial population sizes are low and the fishery
is poorly managed \citep{hilborn_2006}. Both lobster fisheries were
certified by the Marine Stewardship Council \citep{prezramrez_2016-J1}.
Additionally, lobster fisheries are managed via species-specific minimum
catch sizes, seasonal closures, protection of ``berried'' females, and
escapement windows where traps are allowed \citep{dof_website_1993}. It
is uncertain whether such a well-managed fishery will experience
additional benefits from marine reserves. Additionally,
\citet{gelcich_2008} has shown that TURFs alone can have greater biomass
and richness than areas operating under open acces. These increased
attributes perhaps minimize the difference between TURF and reserve.
Further research should focus on evaluating sites in the reserve, TURF,
and open access areas.

While the evaluated reserves have failed to provide fishery benefits up
to now, there are a number of additional ecological, fisheries, and
social benefits. Marine reserves provide protection to a wider range of
species and vulnerable habitat, like coral reefs. These sites can serve
as an insurance against uncertainty and errors in fisheries management,
as well as environmental shocks
\citep{hilborn_2004,hilborn_2006,micheli_2012-EU} (\textbf{,Aalto et al,
in press}). Self-regulation of fishing effort (\emph{i.e.} reduction in
harvest) can serve as a way to compensate for future declines associated
to environmental variation \citep{finkbeiner_2018}. Embarking in a
marine conservation project can bring the community together, which
promotes social cohesion and builds social capital \citep{fulton_2019}.
Furthermore, showing commitment to marine conservation and sustainable
fishing practices allows fishers to have greater bargaining power and
leverage over fisheries management \citep{prezramrez_2012}.

Previous studies have evaluated the potential of implementing marine
reserves in Baja California and connect them to the existing network in
California \citep{arafehdalmau_2017}. Community-based marine reserves in
small-scale fisheries can be helpful conservation and fishery management
tools when appropriately implemented. Lessons learned from these cases
can guide implementation of community-based marine reserves elsewhere.
For the particular case of the marine reserves that we evaluate, the
possibility of expanding reserves or merging existing polygons into
larger areas should be evaluated and proposed to the communities. At the
broader scale, having full community support surely represents an
advantage, but it is important that marine reserves meet essential
design principles such as size and placement. Community-based marine
reserves might have more benefits that result from indirect effects of
the reserves, which should be taken into account when evaluating the
outcomes of similar projects.

\section*{Conflict of Interest Statement}

The authors declare that the research was conducted in the absence of
any commercial or financial relationships that could be construed as a
potential conflict of interest.

\section*{Author Contributions}

JC and EA analyzed and interpreted data, discussed the results, and
wrote the first draft. AHV coordinated fieldwork and collected the data.
AS, AHV, SF, JT, and FM discussed the results and edited the manuscript.

\section*{Funding}

JCVD recevied funding from CONACyT (Beca de Posgrados en el extranjero,
CVU 669403) and the Latin American Fisheries Fellowship Program. AS,
AHV, SF and JT received funding from Marisla Foundation, Packard
Foundation, Walton Family Foundation, Summit Foundation, and Oak
Foundation.

\section*{Acknowledgments}

The authors wish to acknowledge Imelda Amador for contributions on the
governance data, as well as pre-processing biological data. This study
would have not been possible without the effort by members of the
fishing communities here mentioned, who participated in the
data-collection process.

\bibliographystyle{frontiersinSCNS_ENG_HUMS}\bibliography{references}

\section*{Figure captions}

\end{document}