### 2018-09-24 docs/paper/Villasenor-Derbez.Rmd
Finalized manuscript. Waiting for Ere's adscription

---
title: "Effectiveness of community-based marine reserves in small-scale fisheries"
rtitle: "Community-based marine reserves"
author: "Juan Carlos Villaseñor-Derbez$^{1,2,*}$, Eréndira Aceves-Bueno$^{1}$, Álvin Suarez$^{2}$, Stuart Fulton$^{2}$, Arturo Hernández-Velasco$^{2}$, Jorge Torre$^{2}$, Fiorenza Micheli$^{3}$"
institution: "$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
$^{2}$Comunidad y Biodiversidad A.C., Guaymas, Sonora, Mexico\\newline
$^{3}$Hopkins Marine Station and Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA"
corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"
email: "juancarlos@ucsb.edu"
abstract: "Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world. Artisanal fisheries face great challenges since they are difficult to monitor, enforce, and manage. Combining territorial user rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves is believed to improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments. In the last six years, and following a 2012 regulation, 18 TURF-reserves have been implemented in Mexico; their effectiveness has not been formally evaluated. We combine causal inference techniques and a social-ecological systems framework to provide a holistic evaluation of community-based TURF reserves in three coastal communities in Mexico. We find that while reserves have not yet achieved their stated goal of increasing lobster densities, they continue to receive significant support from the fishing communities. A lack of ecological and socioeconomic effects likely results from a combination of factors. First, the lobster fisheries are already well managed, and it is unlikely that reserves might have a detectable effect. Second, some of the reserves are not large enough to protect lobsters' home ranges. Third, some of these reserves might be too young for the effects to show. However, these reserves have shaped small-scale fishers’ way of thinking about marine conservation, which can provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species."
keywords: "TURF-reserves, Causal Inference, Social-Ecological Systems, Marine Protected Areas, Marine Conservation, Small-Scale Fisheries"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    template: template.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

options(knitr.table.format = "latex")
```

```{r load_pkgs}
suppressPackageStartupMessages({
  library(sp)
  library(sf)
  library(ggsn)
  library(tmap)
  library(magrittr)
  library(kableExtra)
  library(broom)
  library(ggrepel)
  library(cowplot)
  library(tidyverse)
})

# Load indicator functions
source(file = here::here("scripts", "abundance.R")) #Calculate abundance per transect
source(file = here::here("scripts", "biomass.R")) # Calculate fish biomass per transect
source(file = here::here("scripts", "trophic.R")) # Calculate mean trophic level per transect

# Load analysis functions
source(file = here::here("scripts", "model.R")) #fits the model
source(file = here::here("scripts", "robust_se.R")) #calculates heteroskedastic-robust Standard Errors
source(file = here::here("scripts", "robust_se_list.R")) #heteroskedastic-robust Standard Errors for stargazer
source(file = here::here("scripts", "plot_model.R")) #plots effect sizes
```

\clearpage

# Introduction

Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. In particular, artisanal fisheries face great challenges since they tend to be hard to monitor and enforce \citep{costello_2012}. Recent research shows that combining Territorial Use Rights for Fisheries (TURFs) with no-take marine reserves (MR) can greatly improve the performance of coastal fisheries and the health of the local resouces \citep{costello_2010-Ix,lester_2017}. Commonly known as TURF-Reserves, these systems increase the benefits of spatial access rights allowing the maintainance of healthy resources \citep{afflerbach_2014-HP,lester_2017}. Although in theory these systems are successful \citep{costello_2010-Ix}, there is little empirical evidence of their effectiveness and the drivers of their success \citep{afflerbach_2014-HP,lester_2017,smallhornwest_2018}.

The performance of these systems depends on how environmental and social factors combine and interact. The science of marine reserves has largely focused on understanding the ecological effects of these areas, which include increased biomass, species richness, and densities of organisms within the protected regions, climate change mitigation, and protection from environmental variability \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}. Modelling studies show that fishery benefits of marine reserves depend on initial stock status and the management under which the fishery operates, as well as reserve size and the amount of larvae exported from these \citep{hilborn_2006,krueck_2017-J1,deleo_2015}. Other research has focused on the relationship between socioeconomic and governance structures and reserve effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. However, to our knowledge, no studies exist that evaluate TURF-reserves from both a social and ecological perspective. This is especially important in social-ecological coastal systems dominated by close interaction and feedbacks between people and natural resources \citep{ostrom_2009-hg}.

TURF-reserves can be created as community-based marine reserves, voluntarily established and enforced by local communities. This bottom-up approach increases compliance and self-enforcement \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}. Community-based spatial closures occur in different contexts, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}. However, without legal recognition no-take regulations are difficult to enforce and fishers rely on the exclusive access granted by the TURF. In an effort to bridge this normative gap, Civil Society Organizations (CSOs) have served as a link between fishers and government, and set out to create a legal framework that solve this governance issue. In Mexico, a new norm was created in 2014 allowing fishers to request the legal recognition of community-based reserves as “Fish Refuge” (*Zona de Refugio Pesquero*; \citet{nom}). Fish refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries. Since 2012, 45 of Fish Refuges have been created along the Pacific, Gulf of California, and Mexican Caribbean coastlines, with 18 of them implemented as TURF-reserves. However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.

Here, we combine causal inference techniques and a social-ecological systems framework to provide a holistic evaluation of community-based marine reserves in three coastal communities in Mexico. These three case studies span a range of ecological and social conditions representative of different regions of Mexico. The objective of this work is twofold. First, to provide a triple bottom line evaluation of the effectiveness of community-based marine reserves that can inform similar processes in other countries. Second, to evaluate the effectiveness of TURF-reserves established as Fish Refuges in Mexico to identify opportunities where improvement or adjustment might lead to increased effectiveness. We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based marine reserves in small-scale fisheries in Mexico and in other regions around the world that are using this tool to manage and rebuild their coastal fisheries.

# Methods

## Study area

We evaluate three TURF-reserves in Mexico (Fig \ref{fig:map}A). The first one was created by the *Buzos y Pescadores de la Baja California* fishing cooperative, located in Isla Natividad in the Baja California Peninsula (Fig \ref{fig:map}B). The main fishery in the island is the spiny lobster (*Panulirus interruptus*), but other resources like finfish, sea cucumber, read sea urchin, snail, and abalone are also an important source of income. In 2006, the community decided to implement two marine reserves within their fishing grounds to protect commercially important invertebrate species; mainly lobster and abalone. While these reserves obtained legal recognition only in 2018, they have been well enforced since their implementation.

The other two TURF-reserves are located in Maria Elena and Punta Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). In contrast with Isla Nativdad, which hosts a well established fishing community, Maria Elena is a fishing camp --visited intermittently during the fishing season-- belonging to the Cozumel fishing cooperative (*SCPP Cozumel*); Punta Herrero is home to the *SCPP José María Azcorra* cooperative, and similar to Isla Natividad hosts a local community. Their main fishery is the Caribbean spiny lobster (*Panulirus argus*), but they also target finfish in the off-season. Maria Elena and Punta Herrero established eight marine reserves in 2012, and four marine reserves in 2013, respectively. All these reserves have been legally recognized as Fishing Refuges since their creation \citep{dof_website_2012,dof_website_2013}.

These communities are representative of their region in terms of ecology, socioeconomic, and governance aspects. Isla Natividad, for example, is part of a greater group of fishing cooperatives belonging to a Federation of Fishing Cooperatives. This group has been identified as a cohesive group that often cooperates to better manage their resources \citep{mccay_2017-1m,mccay_2014-CN,acevesbueno_2017}. Likewise, Maria Elena and Punta Herrero are representative of fishing cooperatives in the Mexican Caribbean, which are also part of a regional Federation. Together, these three communities provide an accurate representation of other fishing communities in each of their regions. While each region has additional communities that have established community-based TURF-reserves, available data would not allow us to perform the in-depth analysis that we undertake. Yet, given the similarities among communities and the socioeconomic and governance setting under which they opperate, it is safe to cautiously generalize our results to other communities in Mexico and other regions around the world.

```{r load_spatial}
# Determine a projection for the maps
proj <- "+proj=longlat +datum=NAD27"

# Load coastline for Mexico and World
data(World)
W <- spTransform(World, proj) %>% 
  st_as_sf()

# Load high-res mexican coastline
coastline_mx <- readRDS(here::here("raw_data", "spatial", "coastline_mx.rds"))

# Load ZRP shapefile
zrp <- st_read(dsn = here::here("raw_data", "spatial"), layer = "ZRP", quiet = T) %>% 
  filter(Name %in% c("El Cabezo",
                     "Gallineros",
                     "Punta Loria",
                     "San Roman Norte",
                     "San Roman Sur",
                     "Anegado de Chal",
                     "Laguna Canche Balam",
                     "Niluc",
                     "Mimis",
                     "El Faro (Pesca de Langosta)",
                     "El Faro")) %>% 
  st_transform(crs = proj) %>% 
  mutate(evaluated = ifelse(Name %in% c("El Cabezo", "El Faro (Pesca de Langosta)", "El Faro"), T, F),
         Comunidad = ifelse(Name %in% c("El Faro", "Laguna Canche Balam", "El Faro (Pesca de Langosta)", "Anegado de Chal"), "Puna Herrero", "Maria Elena"))

# Load Community reserve shapefile
rescom <- st_read(dsn = here::here("raw_data", "spatial"), layer = "Res_Com", quiet = T) %>% 
  filter(Nombre %in% c("La_Plana", "PuntaPrieta")) %>% 
  st_transform(crs = proj)

turf <- st_read(dsn = here::here("raw_data", "spatial"), layer = "turfs", quiet = T) %>% 
  filter(Coop %in% c("SCPP Buzos y Pescadores de la Baja California", "SCPP Cozumel", "SCPP Jose Maria Azcorra"))

```

```{r}
# Define reference map

## Declare points for reference map
location_labels <- data.frame(X = c(-115.24, -87.4611756),
                              Y = c(27.85, 27.85),
                              Location = c("Isla Natividad", "Punta Herrero"),
                              Region = c("Pacific", "Caribbean"),
                              stringsAsFactors = F)

## Create map
mex <- ggplot() +
  geom_sf(data = W) +
  geom_point(data = location_labels,
             aes(x = X, y = Y),
             color = "black",
             fill = "steelblue",
             size = 3,
             shape = 21) +
  theme_classic() +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "none",
        panel.grid.major = element_line(colour = "transparent"),
        axis.text = element_text(size = 7)) +
  lims(x = c(-118.36648, -86.71074), y = c(14.53507, 32.71863)) +
  annotate(geom = "text", x = -102, y = 25, label = "Mexico") +
  annotate(geom = "text", x = -90, y = 25, label = "Gulf of\nMexico") +
  annotate(geom = "text", x = -112, y = 20, label = "Pacific\nOcean") +
  annotate(geom = "text", x = -116.4, y = 27.5, label = "B") +
  annotate(geom = "text", x = -86.8, y = 18.2, label = "C") +
  labs(x = "", y = "")

# Map for SAM

## Declare points for SAM
sam_points <- data.frame(x = c(-87.4646745, -87.539104),
                         y = c(19.3229767, 19.396729),
                         location = c("Punta Herrero", "Maria Elena"))

## Create map
ph <- ggplot() +
  geom_sf(data = turf, fill = "steelblue", alpha = 0.5, size = 1, color = "black") +
  geom_sf(data = coastline_mx) +
  lims(x = c(-87.7, -87.35), y = c(19.15, 19.6)) +
  geom_sf(data = zrp, fill = "red2") +
  theme_classic() +
  theme(panel.grid.major = element_line(colour = "transparent"),
        axis.text = element_text(size = 7)) +
  geom_label_repel(data = sam_points,
                   aes(x = x, y = y, label = location),
                   nudge_x = c(0.1, -0.05),
                   nudge_y = c(-0.05, 0.06)) +
  labs(x = "", y = "") +
  scalebar(data = turf[3,], dist = 5, dd2km = T, model = 'WGS84', location = "bottomright", st.size = 2, st.dist = 0.05, height = 0.1) +
  scale_x_continuous(breaks = seq(-87.7, -87.3, 0.1), limits = c(-87.7, -87.3))

# Map for PBC
pbc <- ggplot() +
  geom_sf(data = turf, fill = "steelblue", alpha = 0.5, size = 1, color = "black") +
  lims(x = c(-115.65, -115.109), y = c(27.7, 28)) +
  geom_sf(data = coastline_mx) +
  geom_sf(data = rescom, fill = "red2") +
  theme_classic() +
  theme(panel.grid.major = element_line(colour = "transparent"),
        axis.text = element_text(size = 7)) +
  labs(x = "", y = "") +
  scalebar(data = turf[4,], dist = 10, dd2km = T, model = 'WGS84', location = "bottomleft", st.size = 2, st.dist = 0.05, height = 0.1)

left <- plot_grid(mex, pbc, ncol = 1, labels = "AUTO")

plot1 <- plot_grid(left, ph, labels = c("", "C"), ncol = 2)
```

## Data collection

```{r load_data}
invert <- read.csv(file = here::here("data", "invertebrates.csv"),
                stringsAsFactors = F)

fish <- read.csv(file = here::here("data", "fish.csv"),
                stringsAsFactors = F)

conapesca <- read.csv(file = here::here("data", "conapesca.csv"),
                stringsAsFactors = F)

gob <- read.csv(here::here("data" ,"governance.csv"),
                strip.white = T,
                stringsAsFactors = F)
```

We use three main sources of information to evaluate these reserves across the ecological, socioeconomic, and governance dimensions. Ecological data come from the annual ecological monitoring of reserve and control areas, carried out by members from each community and personnel from the Mexican CSO *Comunidad y Biodiversidad* ([COBI](www.cobi.org.mx)). Trained divers record richness and abundances of fish and invertebrate species along replicate transects (30x 2 m each) at depths 5-20 m in the reserves and control sites \citep{fulton_2018,fulton_2019}. Size structures are also collected during fish surveys. We define control sites as regions with habitat characteristics similar to the corresponding reserves, and that presumably had a similar probability of being selected as reserves during the design phase. We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve. This provides us with a Before-After-Control-Impact (*i.e.* BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}. BACI designs and causal inference techniques have proven effective to evaluate marine reserves, as they allow us to causally attribute observed changes to the intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites were surveyed annually, and at least once before implementation of the reserves.

Socioeconomic data come from landing receipts reported to the National Commission for Aquaculture and Fisheries (*Comisión Nacional de Acuacultura y Pesca*; CONAPESCA). Data contain monthly lobster landings (Kg) and revenues (MXP) for cooperatives with and without marine reserves. Cooperatives incorporated in this analysis belong to larger regional-level Cooperative Federations, and are exposed to the same markets and institutional frameworks, making them plausible controls \citep{mccay_2017-1m,ayer_2018}. Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \citep{oecd_2017-VV} as:

\begin{equation}
I_t = RI_t\times\frac{CPI_t}{CPI_T}
\label{eqn:cpi}
\end{equation}

Where $I_t$ represents the adjusted income for year $t$ as the product between the reported income for that year and the ratio between the consumer price index in that year ($CPI_t$) to the most recent year’s consumer price index ($CPI_T$).

Data for the operationalization of the social-ecological system were collected at the community--level from official documents used in the creation and designation of the marine reserves \citep{dof_website_2012,dof_website_2013,dof_website_2018} and based on the authors' experience and knowledge of the communities. These include information on the resource system, the resource units, actors, and the governance system itself (Table \ref{table:ses}).

## Data analysis

We evaluate the effect that marine reserves have had on four ecological and two socioeconomic indicators (Table \ref{table:indicators}). Recall that reserves were implemented to protect lobster and other benthic invertebrates. However, we also use the available fish data to test for associated co-benefits.

We use a difference-in-differences analysis to evaluate these indicators. This approach allows us to estimate the effect that the reserve had by comparing trends across time and treatments \citep{moland_2013-VP,Villasenor-Derbez_2018}. The analysis of ecological indicators is performed with a multiple linear regression of the form:

\begin{equation}
I_{itj} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \sigma_jSpp_j + \epsilon
\label{eqn:reg_bio}
\end{equation}

Where year-fixed effects are represented by $\gamma_{it} Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites. The interaction term $\lambda_{it} Year_t\times Zone_i$ represents the mean change in the indicator inside the reserve, for year $t$, with respect to the year of implementation in the control site. When evaluating biomass and densities of the invertebrate or fish communities, we include $\sigma_j$ to control for species-fixed effects.

Socioeconomic indicators are evaluated with a similar approach. Due to data constrains, we only evaluate socioeconomic data for Isla Natividad (2000 - 2014) and Maria Elena (2006 - 2013). Neighboring communities are used as counterfactuals that allow us to control for unobserved time-invariants. Each focal community (Isla Natividad and Maria Elena) has three counterfactual communities.

\begin{equation}
I = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i + \sigma_jCom_j +\epsilon
\label{eqn:soc_reg}
\end{equation}

The model interpretation remains as for Eq \ref{eqn:reg_bio}, but in this case the $Treated$ dummy variable indicates if the community has a reserve ($Treated = 1$) or not ($Treated = 0$) and $\sigma_jCom$ captures community-level fixed-effects. These regression models allow us to establish a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and spatial dynamics \citep{depalma_2018}. The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the treated sites at time $t$ after controlling for other time and space variations (*i.e.* $\gamma_t$ and $\beta$ respectively). All model coefficients were estimated via ordinary least-squares and heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All analyses were performed in R 3.5.0 and R Studio 1.1.453 \citep{R_2018}. Data and code are available on [github.com](https://github.com/jcvdav/ReserveEffect).

We use the social-ecologycal system (SES) framework to evaluate each community as a means of providing an explanation to the biological and socioeconomic results. The use the SES framework standardizes our analysis and allows us to communicate our results in a common language across fields. We based our variable selection primarily on \citet{leslie_2015-na,basurto_2013-oq}, who have previously analyzed Mexican fishing cooperatives using this framework. We also incorporate other relevant information following \citet{difranco_2016-Xw} and \citet{edgar_2014-UO}. Table \ref{table:ses} shows the selected variables, their definition and selected indicators.

```{r model_fit}
# Invertebrates

## Lobsters
L_IN <- invert %>% 
  spread(generoespecie, abundancia, fill = 0) %>% 
  gather(generoespecie, abundancia, -c("comunidad", "ano", "year", "rc", "sitio", "zona", "transecto")) %>%
  filter(generoespecie == "Panulirus interruptus") %>% 
  abundance(com = "Isla Natividad") %>% 
  model(sfe = F, spfe = F)

L_ME <- invert %>% 
  spread(generoespecie, abundancia, fill = 0) %>% 
  gather(generoespecie, abundancia, -c("comunidad", "ano", "year", "rc", "sitio", "zona", "transecto")) %>% 
  filter(generoespecie == "Panulirus argus") %>% 
  abundance(com = "Maria Elena") %>% 
  model(sfe = F, spfe = F)

L_PH <- invert %>% 
  spread(generoespecie, abundancia, fill = 0) %>% 
  gather(generoespecie, abundancia, -c("comunidad", "ano", "year", "rc", "sitio", "zona", "transecto")) %>% 
  filter(generoespecie == "Panulirus argus") %>% 
  abundance(com = "Punta Herrero") %>% 
  model(sfe = F, spfe = F)

## Abundance
Ni_IN <- invert %>% 
  abundance(com = "Isla Natividad") %>% 
  model(sfe = F)

Ni_ME <- invert %>% 
  abundance(com = "Maria Elena") %>% 
  model(sfe = F)

Ni_PH <- invert %>% 
  abundance(com = "Punta Herrero") %>% 
  model(sfe = F)

## Abundance commercially important

# Fish
## Biomass
B_IN <- fish %>% 
  biomass(com = "Isla Natividad") %>% 
  model(sfe = F)

B_ME <- fish%>% 
  biomass(com = "Maria Elena") %>% 
  model(sfe = F)

B_PH <- fish %>% 
  biomass(com = "Punta Herrero") %>% 
  model(sfe = F)

## Abundance
N_IN <- fish %>% 
  abundance(com = "Isla Natividad") %>% 
  model(sfe = F)

N_ME <- fish %>% 
  abundance(com = "Maria Elena") %>% 
  model(sfe = F)

N_PH <- fish %>% 
  abundance(com = "Punta Herrero") %>% 
  model(sfe = F)

```

# Results

The following sections present the effect that marine reserves had on each of the biological and socioeconomic indicators for each coastal community. Results are presented in terms of the difference through time and across sites, relative to the control site on the year of implementation (*i.e.* effect size $\lambda_t$). We also provide an overview of the governance settings of each community, and discuss how these might be related to the effectiveness and performance of the reserves.

## Biological effects

Indicators showed ambiguous responses through time for each reserve. Figure \ref{fig:indicators}A shows positive effect sizes for lobster densities in Isla Natividad and Punta Herrero during the first years, but the effect is eroded through time. In the case of Maria Elena, positive changes were observed in the third and forth year. These effects are in the order of 0.2 extra organisms $\mathrm{m}^{-2}$ for Isla Natividad and Punta Herrero, and 0.01 organisms $\mathrm{m}^{-2}$ for Maria Elena, but are not significantly different from zero ($p > 0.05$). Likewise, no changes were detected in fish biomass or invertebrate and fish densities (\ref{fig:indicators}B-D), where effect sizes oscillated around zero without clear trends. Full tables with model coefficients are presented in the supplementary materials (S1 Table, S2 Table, S3 Table).

## Socioeconomic effects

Lobster landings and revenue were only available for Isla Natividad and Maria Elena (Fig \ref{fig:lobsters}). For all years before implementation, the effect sizes are close to zero, indicating that the control and treatment sites have similar pre-treatment trends, suggesting that these are plausible controls. However, effect sizes do not change after the implementation of the reserve. Interestingly, the negative effect observed for Isla Natividad on year 5 correspond to the 2011 hypoxia events. The only positive change observed in lobster landings is for Isla Natividad in 2014 ($p < 0.1$). The three years of post-implementation data for Maria Elena do not show a significant effect of the reserve. Isla Natividad shows higher revenues after the implementation of the reserve, as compared to the control communities. However, these changes are not significant and are associated to increased variation. Full tables with model coefficients are presented in the supplementary materials (S4 Table, S5 Table).

## Governance

Our analysis of the SES (Table \ref{table:ses}) shows that all analyzed communities share similarities known to foster sustainable resource management and increase reserve effectiveness. For example, fishers opperate within clearly outlined TURFs (RS2, GS6.1.4.3) that provide exclusive access to resources and reserves. Along with their relatively small groups (A1 - Number of relevant actors), Isolation (A3), Operational rules (GS6.2), Social monitoring (GS9.1), and Graduated sanctions (GS10.1), these fisheries have solid governance structures that enable them to monitor their resources and enforce rules to ensure sustainable management. In general, success of conservation initiatives depends on the incentives of local communities to maintain a healthy status of the resources upon which they depend \citep{jupiter_2017}. Due to the clarity of access rights and isolation, the benefits of conservation directly benefit the members of the fishing cooperatives, which have favored the development of efficient community-based enforcement systems. However, our SES analysis also highlights factors that might hinder reserve performance or mask outcomes While total reserve size ranges from 0.2% to 3.7% of the TURF area, individual reserves are often small (RS3), and relatively young (RS5). Additionally, fishers harvest healthy stocks (RS4.1), and it's unlikely that marine reserves will result in increased catches.

```{r}

# Invert plots
## Lobsters
L_plot <- plot_model(model = list(L_IN, L_ME, L_PH),
                     model_names = c("IS", "ME", "PH"))

## Abundance
I_N <- plot_model(model = list(Ni_IN, Ni_ME, Ni_PH),
                  model_names = c("IN", "ME", "PH"))

# Fish plots
## Biomass
F_B <- plot_model(model = list(B_IN, B_ME, B_PH),
                 model_names = c("IS", "ME", "PH"))

## Abundance
F_N <- plot_model(model = list(N_IN, N_ME, N_PH),
                  model_names = c("IN", "ME", "PH")) +
  theme(legend.justification = c(1, 0.1),
        legend.position = c(1, 0),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.direction = "horizontal")

plot2 <- cowplot::plot_grid(plotlist = list(L_plot, F_B, I_N, F_N),
                            labels = "AUTO",
                            ncol = 2)
```

```{r}
conapesca %<>%
   filter(years_centered > -10) %>% 
  mutate(years_centered = as.factor(years_centered),
         year = fct_relevel(years_centered, "0"),
         peso_vivo = peso_vivo/1e3,
         valor = valor/1e6,
         loc = zona,
         zona = ifelse(grupo == "Treated", "Reserva", "Control"),
         grupo = loc)
```

```{r}
C_IN <- conapesca %>% 
  filter(comunidad == "Isla Natividad" | is.na(comunidad)) %>% 
  filter(grupo == "PN") %>%
  lm(peso_vivo ~ zona * year + unidad_economica, data = .)

C_ME <- conapesca %>% 
  filter(comunidad == "Maria Elena" | is.na(comunidad)) %>% 
  filter(grupo == "MC") %>%
  lm(peso_vivo ~ zona * year + unidad_economica, data = .)

V_IN <- conapesca %>% 
  filter(comunidad == "Isla Natividad" | is.na(comunidad)) %>% 
  filter(grupo == "PN") %>%
  lm(valor ~ zona * year + unidad_economica, data = .)

V_ME <- conapesca %>% 
  filter(comunidad == "Maria Elena" | is.na(comunidad)) %>% 
  filter(grupo == "MC") %>%
  lm(valor ~ zona * year + unidad_economica, data = .)

#soc_res se fue

plot_socioeco <- function(model_list, model_names, ylab){

  extract_filter <- function(model, name){
    model %>% 
      robust_se() %>% 
      filter(grepl(pattern = "zonaReserva:year", x = term)) %>% 
      mutate(years_centered = as.numeric(gsub(pattern = "zonaReserva:year",
                                              replacement = "",
                                              x = term)),
             Community = name)
  }

  purrr::map2_df(model_list, model_names, extract_filter) %>% 
    ggplot(aes(x = years_centered, y = estimate, group = Community)) +
    geom_hline(yintercept = 0,
               linetype = "dashed",
               size = 1) +
    geom_vline(xintercept = 0,
               linetype = "dashed",
               size = 1) +
    geom_errorbar(aes(color = Community,
                      ymin = estimate-std.error,
                      ymax = estimate + std.error),
                  width = 0.1,
                  size = 1,
                  position = position_dodge(width = 0.4)) +
    geom_point(aes(color = Community, shape = Community),
               size = 4,
               position = position_dodge(width = 0.4),
               alpha = 0.5) +
    scale_color_brewer(palette = "Set1") +
    labs(x = "Years since implementation", y = ylab) +
    theme(text = element_text(size = 12))
}
```

```{r}

c_plot <- plot_socioeco(model_list = list(C_IN, C_ME),
                        model_names = list("IN", "ME"),
                        ylab = quo(lambda[t]))

v_plot <- plot_socioeco(model_list = list(V_IN, V_ME),
                        model_names = list("IN", "ME"),
                        ylab = quo(lambda[t]))

plot3 <- plot_grid(c_plot, v_plot, ncol = 1, labels = "AUTO")
```

```{r}
save(L_IN, L_ME, L_PH, B_IN, B_ME, B_PH, Ni_IN, Ni_ME, Ni_PH, N_IN, N_ME, N_PH, C_IN, C_ME, V_IN, V_ME, file = here::here("data", "results", "results.RData"))
```

# Discussion

<!-- From Fio: One aspect that is important to discuss is the concurrent effect of environmental variability. In isla Natividad, and the whole Pacifico Norte region, the decade following the establishment of the reserves has seen extreme ENSO events and hypoxia, which have impacted populations above and beyond the effects of fishing and conservation. -->

Our results indicate that these TURF-reserves have not increased lobster densities. Additionally, no co-benefits were identified when using other ecological indicators other than the previously reported buffering effect that reserves can have to environmental variability in Isla Natividad \citep{micheli_2012-EU}. The socioeconomic indicators pertaining landings and revenues showed little to no change after reserve implementation. The coastal ecosystems where these reserves are located have been profoundly affected by climatic and oceanographic extremes, including warming events, extreme storms and prolonged hypoxia \citep{micheli_2012-EU,woodson}. Despite the lack of evidence of the effectiveness of these reserves, most of the communities show a positive perception about their performance and continue to support their presence \citep{ayer_2018}. Understanding the social-ecological context in which these communities and their reserves operate might provide insights to this.

Some works evaluate marine reserves by performing inside-outside \citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD} or before-after comparisons \citep{betti_2017-lq}. The first approach does not address temporal variability, and the second can not distinguish between the temporal trends in a reserve and the entire system \citep{depalma_2018}. Our approach to evaluate the temporal and spatial changes provides a more robust measure of reserve effectiveness. For example, we capture previously described patterns like the rapid increase observed for lobster densities in Isla Natividad on the sixth year (*i.e.* 2012; Fig. \ref{fig:indicators}A) -a year after the hypoxia events described by \citet{micheli_2012-EU}- which caused mass mortality of sedentary organisms such as abalone and sea urchins, but not lobster and finfish. Yet, our empirical approach assumes control sites are a plausible counterfactual for treated sites. This implies that treated sites would have followed the same trend as control sites, had the reserves not been implemented. Nonetheless, temporal trends for each site don't show any significant increases (S1 Table, S2 Table, S3 Table), supporting our findings of lack of change in the indicators used.

A first possible explanation for the lack of effectivenes may be the young age of the reserves. Literature shows that age and enforcement are important factors that influence reserve effectiveness \citep{edgar_2014-UO,babcock_2010}. Isla Natividad has the oldest reserves, and our SES analysis suggests that all communities have a well-established community-based enforcement system. With these characteristics, one would expect the reserves to be effective. Maria Elena and Punta Herrero are relatively young reserves (*i.e.* < 6 years old) and effects may not yet be evident due to the short duration of protection, relative to the life histories of the protected species; other community-based marine reserves in tropical ecosystems may take up to six years to show a spillover effect \citep{dasilva_2015-zX}.

Another key condition for effectiveness is reserve size \citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be attributed to poor ecological coherence in reserve design (*sensu* \citet{rees_2018}. Previous research has shown that reserves in Isla Natividad yield fishery benefits for the abalone fishery \citep{rossetto_2015-V0}. Abalone are less mobile than lobsters, and perhaps the reserves provide enough protection to these sedentary invertebrates, but not lobsters. Design principles developed by \citet{green_2017} for marine reserves in the Caribbean state that reserves "should be more than twice the size of the home range of adults and juveniles", and suggest that reserves seeking to protect spiny lobsters should have at least 14 km across. Furthermore, fishers may favor implementation of reserves that pose low fishing costs due to their small size or location. Our analysis of economic data supports this, as neither landings nor revenues showed the expected short-term costs associated to the first years of reserve implementation \citep{ovando_2016-Wg}.

Even if reserves had appropriate sizes and were placed in optimal locations, there are other plausible explanations for the observed patterns. For instance, marine reserves are only likely to provide fisheries benefits if initial population sizes are low and the fishery is poorly managed \citep{hilborn_2006}. Both lobster fisheries were certified by the Marine Stewardship Council \citep{prezramrez_2016-J1}. Additionally, lobster fisheries are managed via species-specific minimum catch sizes, seasonal closures, protection of "berried" females, and escapement windows where traps are allowed \citep{dof_website_1993}. It is uncertain whether such a well-managed fishery will experience additional benefits from marine reserves. Additionally, \citet{gelcich_2008} have shown that TURFs alone can have greater biomass and richness than areas operating under open acces. These increased attributes perhaps minimize the difference between TURF and reserve, making it difficult to detect such a small difference. Further research should focus on evaluating sites in the reserve, TURF, and open access areas or similar Fish Refuges established without the presence of TURFs where the impact of ther eserves might be larger.

While the evaluated reserves have failed to provide fishery benefits up to now, there are a number of additional ecological, fisheries, and social benefits. Marine reserves provide protection to a wider range of species and vulnerable habitat, like coral reefs. These sites can serve as an insurance against uncertainty and errors in fisheries management, as well as environmental shocks \citep{hilborn_2004,hilborn_2006,micheli_2012-EU,aalto}. Self-regulation of fishing effort (*i.e.* reduction in harvest) can serve as a way to compensate for future declines associated to environmental variation \citep{finkbeiner_2018}. Embarking in a marine conservation project can bring the community together, which promotes social cohesion and builds social capital \citep{fulton_2019}. Showing commitment to marine conservation and sustainable fishing practices allows fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}. Furthermore, the lack of effectiveness observed in these reserves may not be generalizable to other reserves established under the same legal frameowrk (*i.e.* Fish Refges) in Mexico, and future research should aim at evaluating other areas that have also been established as bottom-up processes bt without the presence of TURFs (*e.g.* \citet{dof_websiteC_2012}), or others established through a top-down process (*i.e.* \citet{dof_websiteU_2018}).

Previous studies have evaluated the potential of implementing marine reserves in Baja California and connect them to the existing network in California \citep{arafehdalmau_2017}. Community-based marine reserves in small-scale fisheries can be helpful conservation and fishery management tools when appropriately implemented. Lessons learned from these cases can guide implementation of community-based marine reserves elsewhere. For the particular case of the marine reserves that we evaluate, the possibility of expanding reserves or merging existing polygons into larger areas should be evaluated and proposed to the communities. At the broader scale, having full community support surely represents an advantage, but it is important that marine reserves meet essential design principles such as size and placement. Community-based marine reserves might have more benefits that result from indirect effects of the reserves, which should be taken into account when evaluating the outcomes of similar projects.

\section*{Conflict of Interest Statement}

The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

\section*{Author Contributions}

JC and EA analyzed and interpreted data, discussed the results, and wrote the first draft. AHV coordinated fieldwork and collected the data. AS, AHV, SF, JT, and FM discussed the results and edited the manuscript.

\section*{Funding}

JCVD recevied funding from CONACyT (Beca de Posgrados en el extranjero, CVU 669403) and the Latin American Fisheries Fellowship Program. AS, AHV, SF and JT received funding from Marisla Foundation, Packard Foundation, Walton Family Foundation, Summit Foundation, and Oak Foundation.

\section*{Acknowledgments}
The authors wish to acknowledge Imelda Amador for contributions on the governance data, as well as pre-processing biological data. This study would have not been possible without the effort by members of the fishing communities here mentioned, who participated in the data-collection process.

\clearpage

\bibliographystyle{frontiersinSCNS_ENG_HUMS}
\bibliography{references}

\clearpage

\section*{Figure captions}

```{r, fig.cap = "\\label{fig:map}Location of the three coastal communities studied (A). Isla Natividad (B) is located off the Baja California Peninsula, Maria Elena and Punta Herrero (C) are located in the Yucatan Peninsula. Blue polygons represent the TURFs, and red polygons the marine reserves."}
plot1
```

```{r, fig.width = 6.5, fig.height = 4, fig.cap = "\\label{fig:indicators}Effect sizes for marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares) for lobster densities (*Panulirus spp*; A), fish biomass (B), invertebrate densities (C), and fish densities (D). Plots are ordered by survey type (left column: invertebrates; right column: fish). Points are jittered hotizontally to avoid overplotting. Points indicate the effect size and standard errors. Years have been centered to year of implementation."}
plot2
```

```{r, fig.width = 5, fig.height = 4, fig.cap = "\\label{fig:lobsters}Effect sizes for lobster catches (A) and revenues (B) in at Isla Natividad (IN; red circles) and Maria Elena (ME; blue triangles). Points indicate the effect size and standard errors. Years have been centered to year of implementation."}
plot3
```

```{r}
tibble(Category = c("Biological", "Biological","Biological","Biological", "Socioeconomic", "Socioeconomic"),
           Indicator = c("Lobster density", "Invertebrate density", "Fish biomass", "Fish density", "Income from target species", "Landings from target species"),
       Units = c("org $\\mathrm{m}^{-2}$", "org $\\mathrm{m}^{-2}$", "Kg $\\mathrm{m}^{-2}$", "org $\\mathrm{m}^{-2}$", "M MXP", "Metric Tonnes")) %>% 
  knitr::kable(caption = "\\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.", format = "latex", escape = F) %>% 
  kable_styling(latex_options = "HOLD_position")
```

```{r}
read.csv(here::here("raw_data", "ses.csv")) %>% 
  select(-Dim) %>% 
  knitr::kable(caption = "\\label{table:ses}Variables for the Social-Ecological System analysis (IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero). Alphanumeric codes follow \\citet{basurto_2013-oq}; an asterisk (*) denotes variables incorporated based on \\citet{difranco_2016-Xw} and \\citet{edgar_2014-UO}.",
               format = "latex",
               escape = F) %>% 
  kable_styling(latex_options = c("HOLD_position", "scale_down"),
                font_size = 11) %>% 
  column_spec(column = 1, width = "6.5cm") %>% 
  column_spec(column = 2, width = "12cm") %>% 
  group_rows(group_label = "Resource System (RS)", start_row = 1, end_row = 5) %>% 
  group_rows(group_label = "Resource Unit (RU)", start_row = 6, end_row = 6) %>% 
  group_rows(group_label = "Actors (A)", start_row = 7, end_row = 8) %>% 
  group_rows(group_label = "Governance system (G)", start_row = 9, end_row = 12)
```