### 2018-12-03 docs/test.Rmd
Added packrat libraries and docs helpers

---
title: "Effectiveness of community-based TURF-reserves in small-scale fisheries"
rtitle: "Community-based marine reserves"
author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,2}$, Stuart Fulton$^{3}$, Álvin Suarez$^{3}$, Arturo Hernández-Velasco$^{3}$, Jorge Torre$^{3}$, Fiorenza Micheli$^{4}$"
institution: "$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
$^{2}$Nicholas School of the Environment, Duke University, Beaufort, NC, USA\\newline
$^{3}$Comunidad y Biodiversidad A.C., Guaymas, Sonora, Mexico\\newline
$^{4}$Hopkins Marine Station and Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA"
corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"
email: "juancarlos@ucsb.edu"
abstract: "Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world. Small-scale fisheries face great challenges since they are difficult to monitor, enforce, and manage. Combining territorial user rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves can improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments. In the last 12 years, 18 old and new community-based TURF-reserves gained legal recognition thanks to a 2014 regulation; their effectiveness has not been formally evaluated. We combine causal inference techniques and the Social-Ecological Systems framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico. We find that while reserves have not yet achieved their stated goal of increasing the density of lobster and other benthic invertebrates, they continue to receive significant support from the fishing communities. A lack of clear ecological and socioeconomic effects likely results from a combination of factors. First, local fisheries are already well managed, and it is unlikely that reserves might have a detectable effect. Second, some of the reserves are not large enough to protect mobile species, like lobster. Third, some of these reserves might be too young for the effects to show. Fourth, variable and extreme oceanographic conditions have impacted harvested populations. However, these reserves have shaped small-scale fishers’ way of thinking about marine conservation, which can provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species."
keywords: "TURF-reserves, Causal Inference, Social-Ecological Systems, Marine Protected Areas, Marine Conservation, Small-Scale Fisheries"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    template: docs/template.tex
---