### 2018-12-06 scripts/data_cleaning_logbook.Rmd
Added missing zero's for fish and inverts. This fixes #48

---
title: "Data Cleaning logbook"
author: "Juan Carlos Villaseñor-Derbez"
date: "last compilation: `r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages({
  library(rfishbase)
  library(tidyverse)
})
```

# Ecological data

## Invertebrates

```{r}
invert <- read.csv(here::here("raw_data", "invert_data.csv"), strip.white = T, stringsAsFactors = F) %>% 
  janitor::clean_names() %>% 
  rename(generoespecie = genero_especie) %>% 
  filter(comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  filter(!is.na(rc),
         rc %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
                   "Cabezo - Cabezo (Control)",
                   "Manchon - Manchon (Control)",
                   "El Faro - El Faro (Control)")) %>% 
  filter(!(rc %in% c("El Faro - El Faro (Control)") & ano == 2012)) %>% 
  filter(between(abundancia, 0.1, 100)) %>% 
  group_by(comunidad, ano, rc, sitio, zona, transecto, generoespecie) %>% 
  summarize(abundancia = sum(abundancia)) %>% 
  ungroup() %>% 
  mutate(year = case_when(comunidad == "Isla Natividad" ~ ano - 2006,
                         comunidad == "Maria Elena" ~ ano - 2012,
                         TRUE ~ ano - 2013)) %>% 
  spread(generoespecie, abundancia, fill = 0) %>% 
  gather(generoespecie, abundancia, -c("comunidad", "ano", "year", "rc", "sitio", "zona", "transecto")) %>% 
  select(comunidad, ano, year, rc, sitio, zona, transecto, generoespecie, abundancia) %>% 
  arrange()
```

## Fish

### Species biology

```{r}
species_bio <- MPAtools::species_bio %>% 
  mutate(Genus = as.character(str_split(GeneroEspecie, pattern = " ", simplify = T)[,1]),
         Genus = str_replace(Genus, " ", ""))

needed_spp <- c("Kyphosus incisor",
                "Kyphosus sectatrix",
                "Pterois volitans",
                "Scarus guacamaia",
                "Chaetodon capistratus",
                "Chaetodon capistratus",
                "Chaetodon striatus",
                "Calamus pennatula")

FB_update <- rfishbase::length_weight(needed_spp, fields = c("a", "b")) %>% 
  select(GeneroEspecie = sciname, a, b) %>% 
  mutate(GeneroEspecie = case_when(GeneroEspecie %in% c("Kyphosus sectatrix", "Kyphosus incisor") ~ "Kyphosus",
                                   GeneroEspecie %in% c("Chaetodon capistratus", "Chaetodon striatus") ~ "Chaetodon",
                                   TRUE ~ GeneroEspecie)) %>% 
  group_by(GeneroEspecie) %>% 
  summarize_all(mean, na.rm = T) %>% 
  ungroup()

needed_genus <- c("Mycteroperca",
                  "Chromis",
                  "Stegastes",
                  "Haemulon",
                  "Acanthurus",
                  "Sebastes",
                  "Lutjanus",
                  "Scarus")

GenusLevelBio <- species_bio %>% 
  filter(Genus %in% needed_genus) %>%
  select(GeneroEspecie = Genus, a, b) %>% 
  group_by(GeneroEspecie) %>% 
  summarize_all(mean, na.rm = T) %>% 
  ungroup()

species_bio %<>%
  select(GeneroEspecie, a, b) %>% 
  rbind(FB_update) %>% 
  rbind(GenusLevelBio) %>% 
  janitor::clean_names() %>% 
    rename(generoespecie = genero_especie)
```

### Clean data

```{r}
fish <- read.csv(here::here("raw_data", "fish_data.csv"), strip.white = T, stringsAsFactors = F) %>% 
  janitor::clean_names() %>% 
  rename(generoespecie = genero_especie) %>% 
  filter(comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  filter(!is.na(rc),
         rc %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
                   "Cabezo - Cabezo (Control)",
                   "Manchon - Manchon (Control)",
                   "El Faro - El Faro (Control)")) %>% 
  filter(!(rc %in% c("El Faro - El Faro (Control)") & ano == 2012)) %>% 
  filter(abundancia > 0) %>% 
  mutate(generoespecie = case_when(generoespecie == "Cephalopholis cruentatus" ~ "Cephalopholis cruentata",
                                   generoespecie == "Haemulon macrostomun" ~ "Haemulon macrostomum",
                                   generoespecie == "Xcochin spp" ~ "Balistes capriscus",
                                   generoespecie == "Pterois volitans/miles" ~ "Pterois volitans",
                                   generoespecie == "Mycteroperca spp" ~ "Mycteroperca",
                                   generoespecie == "Chromis spp" ~ "Chromis",
                                   generoespecie == "Damisela spp" ~ "Stegastes",
                                   generoespecie == "Haemulon spp" ~ "Haemulon",
                                   generoespecie == "Lancero spp" ~ "Acanthurus",
                                   generoespecie == "Kyphosus sectatrix/incisor" ~ "Kyphosus",
                                   generoespecie == "Sebastes spp" ~ "Sebastes",
                                   generoespecie == "Loro spp" ~ "Scarus",
                                   generoespecie == "Lutjanus spp" ~ "Lutjanus",
                                   generoespecie == "Mariposa spp" ~ "Chaetodon",
                                   generoespecie == "Pluma spp" ~ "Calamus pennatula",
                                   generoespecie == "Sparisoma aurolineatum" ~ "Sparisoma aurofrenatum",
                                   generoespecie == "Vieja spp" ~ "Bodianus rufus",
                                   TRUE ~ generoespecie)) %>% 
  left_join(species_bio, by = c("generoespecie")) %>% 
  group_by(comunidad, ano, rc, sitio, zona, transecto, generoespecie, talla, a,b) %>% 
  summarize(abundancia = sum(abundancia)) %>% 
  ungroup() %>% 
  mutate(year = case_when(comunidad == "Isla Natividad" ~ ano - 2006,
                         comunidad == "Maria Elena" ~ ano - 2012,
                         TRUE ~ ano - 2013)) %>% 
  spread(generoespecie, abundancia, fill = 0) %>% 
  gather(generoespecie, abundancia, -c("comunidad", "ano", "year", "rc", "sitio", "zona", "transecto", "talla", "a", "b")) %>% 
  select(comunidad, ano, year, rc, sitio, zona, transecto, generoespecie, talla, abundancia, a, b) %>% 
  arrange() %>% 
  drop_na(a, b)
```

# Socioeconomic data

## CPI

```{r}
cpi <- read.csv(here::here("raw_data", "CPI.csv"), stringsAsFactors = F, strip.white = T)

cpi_2014 <- cpi$CPI[cpi$Ano == 2014]
```

## Landings and revenue

```{r}
cooperatives_t <- c("Scpp cozumel scl",
                    "Scpp jose maria azcorra scl",
                    "Scpp buzos y pescadores de la baja california scl")

cooperatives_c <- c("Scpp langosteros del caribe scl",
                    "Scpp vigia chico scl",
                    "Scpp pesc nacionales de abulon sc de rl",
                    "Scpp la purisima scl",
                    "Scpp bahia tortugas scl")

cooperatives <- c(cooperatives_t, cooperatives_c)
```

```{r}
L_conapesca <- tibble(UnidadEconomica = c("buzos y pescadores de la baja california",
                                          "bahia tortugas",
                                          "pesc nacionales de abulon",
                                          "la purisima",
                                          "cozumel",
                                          "jose maria azcorra",
                                          "vigia chico",
                                          "langosteros del caribe"),
                      Comunidad = c("Isla Natividad",
                                    NA, NA, NA, 
                                    "Maria Elena",
                                    "Punta Herrero",
                                    NA, NA),
                      Zona = c("PN", "PN", "PN", "PN", "MC", "MC", "MC", "MC"),
                      Start = c(2006, 2006, 2006, 2006, 2012, 2013, 2012, 2012))
```

```{r}
conapesca <- readRDS(here::here("raw_data","conapesca.rds")) %>% 
  mutate(UnidadEconomica = as.character(UnidadEconomica),
         LugarDeCaptura = as.character(LugarDeCaptura)) %>% 
  mutate(UnidadEconomica = ifelse(UnidadEconomica == "Scpp bahia tortugas s.c de r.l", 
                                  "Scpp bahia tortugas scl",
                                  UnidadEconomica)) %>% 
  filter(UnidadEconomica %in% cooperatives) %>%
  filter(NombreComun %in% c("Langosta roja ent. fca.", "Langosta ent. fca.")) %>%
  filter(!(UnidadEconomica == "Scpp cozumel scl" & !LugarDeCaptura %in% c("Bahia espiritu santo",
                                                                          "Bahia del del espiritu santo",
                                                                          "Bahia espiritu santo"))) %>% 
  group_by(UnidadEconomica, Ano, NombreComun) %>%
  summarize(PesoDesembarcado = sum(PesoDesembarcado), 
            PesoVivo = sum(PesoVivo),
            Valor = sum(Valor)) %>%
  ungroup() %>% 
  filter(!(UnidadEconomica == "Scpp buzos y pescadores de la baja california scl" & NombreComun == "Langosta ent. fca.")) %>% 
  left_join(cpi, by = "Ano") %>% 
  mutate(Valor = Valor * (CPI / cpi_2014),
         Group = case_when(UnidadEconomica %in% cooperatives_t ~ "Treated",
                           TRUE ~ "Control"),
         UnidadEconomica = gsub(pattern = "Scpp ", replacement = "", x = UnidadEconomica),
         UnidadEconomica = gsub(pattern = " scl", replacement = "", x = UnidadEconomica),
         UnidadEconomica = gsub(pattern = " s.c de r.l", replacement = "", x = UnidadEconomica),
         UnidadEconomica = gsub(pattern = " sc de rl", replacement = "", x = UnidadEconomica)) %>% 
  left_join(L_conapesca, by = "UnidadEconomica") %>% 
  mutate(Post = ifelse(Ano <= Start, 0, 1),
         years_centered = Ano - Start) %>% 
  select(Zona, Grupo = Group, Comunidad, UnidadEconomica, Ano, Post, years_centered, NombreComun, PesoVivo, PesoDesembarcado, Valor, CPI) %>% 
  janitor::clean_names()
```

# Export all data

```{r}
write.csv(x = invert, file = here::here("data", "invertebrates.csv"), row.names = F)

write.csv(x = fish, file = here::here("data", "fish.csv"), row.names = F)

write.csv(x = conapesca, file = here::here("data", "conapesca.csv"), row.names = F)

```