### 2018-12-27 docs/manuscript.tex

\documentclass{frontiersSCNS}
\usepackage{url,hyperref,lineno,microtype,subcaption}
\usepackage[onehalfspacing]{setspace}
\usepackage{float}

\linenumbers

\usepackage[utf8]{inputenc}
\floatplacement{figure}{H}

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira
Aceves-Bueno\(^{1,2}\), Stuart Fulton\(^{3}\), Álvin Suarez\(^{3}\),
Arturo Hernández-Velasco\(^{3}\), Jorge Torre\(^{3}\), Fiorenza
Micheli\(^{4}\)}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
of California, Santa Barbara, Santa Barbara, CA, USA \newline
\(^{2}\)Nicholas School of the Environment, Duke University, Beaufort,
NC, USA \newline \(^{3}\)Comunidad y Biodiversidad A.C., Guaymas,
Sonora, Mexico \newline \(^{4}\)Hopkins Marine Station and Center for
Ocean Solutions, Stanford University, Pacific Grove, CA, USA}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California,
Santa Barbara, Santa Barbara, CA, 93106}

\def\corrEmail{\href{mailto:juancarlos@ucsb.edu}{\nolinkurl{juancarlos@ucsb.edu}}}

\begin{document}
\onecolumn
\firstpage{1}

\title[Community-based TURF-reserves]{Effectiveness of community-based TURF-reserves in small-scale fisheries}

\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}

\maketitle

\begin{abstract}

Coastal marine ecosystems provide livelihoods for small-scale fishers
and coastal communities around the world. Small-scale fisheries face
great challenges since they are difficult to monitor, enforce, and
manage. Combining territorial user rights for fisheries (TURF) with
no-take marine reserves to create TURF-reserves can improve the
performance of small-scale fisheries by buffering fisheries from
environmental variability and management errors, while ensuring that
fishers reap the benefits of conservation investments. In the last 12
years, 18 old and new community-based Mexican TURF-reserves gained legal
recognition thanks to a 2014 regulation; their effectiveness has not
been formally evaluated. We combine causal inference techniques and the
Social-Ecological Systems framework to provide a holistic evaluation of
community-based TURF-reserves in three coastal communities in Mexico. We
find that while reserves have not yet achieved their stated goal of
increasing the density of lobster and other benthic invertebrates, they
continue to receive support from the fishing communities. A lack of
clear ecological and socioeconomic effects likely results from a
combination of factors. First, some of these reserves might be too young
for the effects to show. Second, the reserves are not large enough to
protect mobile species, like lobster. Third, variable and extreme
oceanographic conditions have impacted harvested populations. Fourth,
local fisheries are already well managed, and it is unlikely that
reserves might have a detectable effect in catches. However, these
reserves may provide a foundation for establishing additional, larger
marine reserves needed to effectively conserve mobile species.

\medskip
\tiny
 \keyFont{ \section{Keywords:} TURF-reserves, Causal Inference, Social-Ecological Systems, Marine
Protected Areas, Marine Conservation, Small-Scale Fisheries}

\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Marine ecosystems around the world sustain significant impacts due to
overfishing and unsustainable fishing practices
\citep{pauly_2005-qV,worm_2006-IB,halpern_2008-dK}. In particular,
small-scale fisheries face great challenges since they tend to be hard
to monitor and enforce \citep{costello_2012}. One of the many approaches
taken to improve the performance of coastal fisheries and health of the
local resources is through the implementation of Territorial Use Rights
for Fisheries (TURFs) that contain no-take marine reserves within them,
thus creating TURF-reserve systems
\citep{afflerbach_2014,gelcich_2015,lester_2017}.

TURFs are a fisheries management tool in which a well defined group of
fishers have exclusive access to an explicitly delimited portion of the
ocean. They promote a sense of stewardship and incentivise resource
users to sustainably manage their resources
\citep{gelcich_2008,costello_2010,mccay_2014}. On the other hand,
no-take marine reserves (marine reserves from hereinafter) are areas
where all extractive activities are off-limits. These can be implemented
to protect biodiversity but also as fishery management tools that
restrict fishing effort and gears and therefore aid in the recovery of
marine stocks. These instruments can be combined by establishing a
marine reserve within a TURF, thus making them TURF-reserves
\citep{afflerbach_2014,gelcich_2015,lester_2017}.

Conservation science has shown how marine reserves lead to increased
biomass, species richness, and abundance within the protected regions
\citep{lester_2009}, and that these may have a series of additional
benefits like climate change mitigation, protection from environmental
variability, and fisheries benefits
\citep{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. Likewise,
research on TURFs has shown that these areas have higher abundance of
targeted species than sites operating under open access and even similar
to that of marine reserves \citep{gelcich_2008,gelcich_2012}. The
benefits resulting from reserves established within TURFs (\emph{i.e.}
TURF-reserves) should be captured exclusively by the group of fishers
with exclusive access \citep{gelcich_2015}. Although in theory these
systems are successful \citep{smallhornwest_2018}, there is little
empirical evidence of their effectiveness and the drivers of their
success.

TURF-reserve systems are inherently intricate social-ecological systems,
and their effectiveness must depend on how environmental and social
factors combine and interact \citep{gelcich_2015}. This is especially
important in social-ecological coastal systems dominated by close
interactions and feedbacks between people and natural resources
\citep{ostrom_2009}. There is a growing body of literature focusing on
the relations between socioeconomic and governance structures and
reserve effectiveness
\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_,bergseth_2018}.
However, to our knowledge, no studies exist that evaluate TURF-reserves
from both a social and ecological perspective.

Recent norms in fisheries regulation in Mexico provide a ripe
opportunity to study the effectiveness of community-based TURF-reserves
in small-scale fisheries. In Mexico, a norm created in 2014 allows
fishers to request legal recognition of community-based reserves as
``Fish Refuges'' (\emph{Zona de Refugio Pesquero}; \citet{nom}). Since
2012, old and new marine reserves have gained legal recognition as
Fishing Refuges. Of these, 18 were originally implemented as within
TURFs. However, their effectiveness has not yet been formally evaluated
and reported in the scientific literature.

Here, we combine causal inference techniques and the Social-Ecological
Systems (SES) framework to evaluate community-based TURF-reserves in
three coastal communities in Mexico. These three case studies span a
range of ecological and social conditions representative of different
regions of Mexico. The objective of this work is twofold. First, to
provide a holistic evaluation of the effectiveness of community-based
TURF-reserves in terms of the changes in biological and socioeconomic
indicators and the governance settings under which these develop, which
may inform similar processes in other countries. Second, to identify
opportunities where improvement or adjustment might lead to increased
effectiveness. We draw from lessons learned in these three case studies
and provide management recommendations to maximize the effectiveness of
community-based TURF-reserves in small-scale fisheries where this tool
is used to manage and rebuild their coastal fisheries.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{turf-reserves-in-mexico}{%
\subsection{TURF-reserves in Mexico}\label{turf-reserves-in-mexico}}

Before discussing our data collection methods and describing our
analyses, our case studies warrant some more background. Community-based
marine reserves that are implemented within TURFs are a form of
TURF-reserves, voluntarily established and enforced by local
communities. This bottom-up approach increases compliance and
self-enforcement, and reserves can yield benefits similar to
systematically-designed reserves \citep{beger_2004,smallhornwest_2018}.
Community-based spatial closures occur in different contexts, like the
\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
\citep{bohnsack_2004,johannes_2002}. However, community-based reserves
can be hard to enforce if they are not legally recognized. In such
conditions, TURF fishers must rely on the exclusive access of the TURF
to maintain high levels of compliance.

In an effort to bridge this normative gap, Mexican Civil Society
Organizations (CSOs) served as a link between fishers and government,
and created a legal framework that solves this governance issue
(\emph{i.e.} Fish Refuges; \citet{nom}). Fish Refuges can be implemented
as temporal or partial reserves, which can protect one, some, or all
resources within their boundaries. One of the ways in which fishing
communities have taken advantage of this new tool is by implementing
temporal marine reserves within their TURFs with a defined expiration
date (often 5 years). After these five years have passed, fishers have
the opportunity to open the reserves to fishing or continue to have
them. Our work focuses on Fish Refuges implemented as community-based
TURF-reserves that occur in small-scale fisheries.

The common setup of community-based TURF-reserves in Mexico is the
following. Fishers from a given community are assembled in fishing
cooperatives which have exclusive fishing rights over a spatially
delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig
\ref{fig:map}A). Each TURF is elusively fished by one cooperative, and
each community usually hosts no more than one cooperative. The profits
from each TURF are shared amongst all fishers from the cooperative.
Fishing cooperatives interested in implementing marine reserves work
with CSOs to implement marine reserves within their TURFs (\emph{i.e.}
TURF-reserves). Fishers then ask the government to grant legal
recognition to their TURF-reserves as Fish Refuges following a series of
studies outlined in the regulation \citep{nom}.

\hypertarget{study-areas}{%
\subsection{Study areas}\label{study-areas}}

We evaluate three community-based no-take TURF-reserves implemented in
Mexican TURF-managed fisheries, therefore making them TURF-reserves (Fig
\ref{fig:map}A). The first one was created by the \emph{Buzos y
Pescadores de la Baja California} fishing cooperative, located in Isla
Natividad in the Baja California Peninsula (Fig \ref{fig:map}B). The
main fishery in the island is the spiny lobster (\emph{Panulirus
interruptus}), but other resources like finfish, sea cucumber, read sea
urchin, snail, and abalone are also an important source of income. In
2006, the community decided to implement two marine reserves within
their fishing grounds to protect commercially important invertebrate
species; mainly lobster and abalone. These reserves obtained legal
recognition in 2018 \citep{dof_website_2018}.

The other two TURF-reserves are located in Maria Elena and Punta
Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). In contrast with
Isla Nativdad, which hosts a well established fishing community, Maria
Elena is a fishing camp --visited intermittently during the fishing
season-- belonging to the \emph{Cozumel} fishing cooperative; Punta
Herrero is home to the \emph{José María Azcorra} fishing cooperative,
and similar to Isla Natividad hosts a local community. Their main
fishery is the Caribbean spiny lobster (\emph{Panulirus argus}), but
they also target finfish in the off-season. Maria Elena and Punta
Herrero established eight and four marine reserves in 2012 and 2013,
respectively. These reserves have been legally recognized as Fishing
Refuges since their original implementation
\citep{dof_website_2012,dof_website_2013} and re-enactments
\citep{dof_website_2017b}.

These communities are representative of their region in terms of
ecology, socioeconomic, and governance aspects. Isla Natividad, for
example, is part of a greater group of fishing cooperatives belonging to
a Federation of Fishing Cooperatives. This group has been identified as
a cohesive group that cooperates to better manage their resources
\citep{mccay_2014,mccay_2017,acevesbueno_2017}. Likewise, Maria Elena
and Punta Herrero are representative of fishing cooperatives in the
Mexican Caribbean, which are also part of a regional Federation.
Together, these three communities provide an accurate representation of
other fishing communities that have been historically manged with TURFs
in each of their regions. While each region has additional communities
that have established community-based TURF-reserves, available data
would not allow us to perform the in-depth causal inference analysis
that we undertake. Yet, given the similarities among communities and the
socioeconomic and governance setting under which they operate, it is
safe to cautiously generalize our insights to other similar
community-based TURF-reserves in Mexico and elsewhere.

The regulation governing the implementation of Fish Refuges states that
these are fishery management tools intended to have biological or
socioeconomic benefits \citep{nom}. For this reason, the main portion of
our analyses focuses on a series of biological and socioeconomic
indicators that may respond to reserve implementation. However, the
effectiveness of conservation and fisheries management interventions
also depends on the social and governance structures in place. We
therefore incorporate a reduced version of the Social Ecological Systems
framework \citep{ostrom_2009} and evaluate variables and indicators
known to aid and hinder the effectiveness of management interventions in
conservation and fisheries. The incorporation of the SES is not intended
to relate different levels of governance with reserve effectiveness, but
rather help provide context on the social-ecological system in which
reserves develop. The following two sections describe our data
collection methods and analyses of biological and socioeconomic
indicators as well as the SES analysis.

\hypertarget{data-collection}{%
\subsection{Data collection}\label{data-collection}}

We use three main sources of information to evaluate these reserves
across the ecological, socioeconomic, and governance dimensions.
Ecological data come from the annual ecological monitoring of reserve
and control sites. Reserve sites are those within the reserves, and thus
no fishing takes place. Control sites are areas that meet the following
criteria: i) habitat characteristics are similar to the corresponding
reserves, ii) presumably had a similar probability of being selected as
reserves during the design phase, iii) are located within the TURF,
where fishing occurs, and iv) Are not directly adjacent to the reserves.
We focus our evaluation on sites where data are available for reserve
and control sites, before and after the implementation of the reserve.
This provides us with a Before-After-Control-Impact (\emph{i.e.} BACI)
sampling design that allows us to capture and control for temporal and
spatial dynamics \citep{stewartoaten_1986,depalma_2018} and causally
attribute the changes to the reserve
\citep{francinifilho_2008,moland_2013,Villasenor-Derbez_2018}.

The biological data are collected by members from each community and
personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
(\href{www.cobi.org.mx}{COBI}). Trained divers record species richness
and abundances of fish and invertebrate species along replicate
transects (30 \(\times\) 2 m each) at depths 5-20 m in the reserves and
control sites \citep{suman_2010-ez,fulton_2018,fulton_2019}. Size
structures are also collected during fish surveys. All sites were
surveyed annually, and at least once before implementation of the
reserves. A summary of sampling effort is shown in the supplementary
materials (Tables S1-S2).

Socioeconomic data come from landing receipts reported to the National
Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de
Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings
(Kg) and revenues (MXP) for TURF-managed cooperatives with and without
marine reserves. In this case our treated unit are the cooperatives
(\emph{i.e.} communities) that have implemented a reserve within their
TURF, and the controls are nearby communities that have a TURF but did
not implement a reserve. Cooperatives incorporated in this analysis
belong to larger regional-level Cooperative Federations, and are exposed
to the same markets and institutional frameworks, making them plausible
controls \citep{mccay_2017,ayer_2018}. Landings and revenues were
aggregated at the cooperative-year level, and revenues were adjusted to
represent 2014 values by the Consumer Price Index for Mexico
\citep{oecd_2017}. A table with summary statistics for this data is
provided in the supplementary materials \textbf{(Table S3, Figure S5)}.

Data for the evaluation of the SES were collected at the community-level
from official documents used in the creation and designation of the
marine reserves. These includes the technical studies that the
cooperatives submit when they request recognition of their reserves, as
well as the official enactments
\citep{dof_website_2012,dof_website_2013,dof_website_2018}. We also
complimented information based on the authors' experience and knowledge
of the communities. We collected information on the Resource Systems,
Resource Units, Actors, and Governance System (Table \ref{table:ses}).
The next sections further describe the variables and indicators chosen
for each of these.

\hypertarget{data-analysis}{%
\subsection{Data analysis}\label{data-analysis}}

We evaluate the effect that the TURF-reserves have had on four
ecological and two socioeconomic indicators shown in Table
\ref{table:indicators}. Recall that reserves were implemented to protect
lobster and other benthic invertebrates. However, we also use the
available fish and invertebrate data to test for associated co-benefits.

We use a difference-in-differences analysis to evaluate these
indicators. This approach is widely used in econometric literature to
estimate the average treatment effect of an intervention, like the
impact of minimum wage increases on employment rates \citep{card_1994}.
In our case it allows us to estimate the effect that the reserve had on
each biological and socioeconomic indicator (Table
\ref{table:indicators}) by comparing trends across time and treatments
\citep{moland_2013,Villasenor-Derbez_2018}. To perform
difference-in-differences, we regress the indicator of interest on a
dummy variable for treatment, a dummy variable for years, and the
interaction term between these with a multiple linear regression of the
form:

\begin{equation}
I_{i,t,j} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \epsilon_{i,t,j}
\label{eqn:reg_bio}
\end{equation}

Where year-level fixed effects capturing a temporal trend are
represented by \(\gamma_t Year_t\), and \(\beta Zone_i\) captures the
difference between reserve (\(Zone = 1\)) and control (\(Zone = 0\))
sites. The effect of the reserve is captured by the \(\lambda_t\)
coefficient, and represents the difference observed between the control
site before the implementation of the reserve and the treated sites at
time \(t\) after controlling for other time and space variations
(\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively). Therefore, we
would expect this term to be positive if the indicator increases because
of the reserve implementation. Finally, \(\epsilon_{i,t,j}\) represents
the error term of the regression.

Socioeconomic indicators are evaluated with a similar approach. Due to
data constrains, we only evaluate socioeconomic data for Isla Natividad
(2000 - 2014) and Maria Elena (2006 - 2013). Neighboring communities are
used as counterfactuals that allow us to control for unobserved
time-invariants. Each focal community (Isla Natividad and Maria Elena)
has three counterfactual communities.

\begin{equation}
I_{i,t,j} = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i +\epsilon_{i,t,j}
\label{eqn:soc_reg}
\end{equation}

The coefficient interpretations remains as for Eq. \ref{eqn:reg_bio},
but in this case the \(Treated\) dummy variable indicates if the
community has a reserve (\(Treated = 1\)) or not (\(Treated = 0\)).
These regression models allow us to establish a causal link between the
implementation of marine reserves and the observed trends by accounting
for temporal and site-specific dynamics \citep{depalma_2018}. All model
coefficients were estimated via ordinary least-squares and
heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All
analyses were performed in R version 3.5.1 and R Studio version 1.1.456
\citep{R_2018}. All data and code needed to reproduce our analyses are
available in a \href{https://github.com/jcvdav/ReserveEffect}{GitHub
repository}.

We use the SES framework to evaluate each community and create a
narrative that provides context for each community. The use of this
framework standardizes our analysis and allows us to communicate our
results in a common language across fields by using a set of previously
defined variables and indicators. We based our variable selection
primarily on \citet{leslie_2015-na} and \citet{basurto_2013-oq}, who
operationalized and analyzed Mexican fishing cooperatives using this
framework, and identified the key variables relevant to fishing
cooperatives in Mexico. We also incorporate other relevant variables
known to influence reserve performance following
\citet{difranco_2016-Xw} and \citet{edgar_2014-UO}. Table
\ref{table:ses} shows the selected variables, along with definitions and
values.

\hypertarget{results}{%
\section{Results}\label{results}}

The following sections present the effect that marine reserves had on
each of the biological and socioeconomic indicators for each coastal
community. Results are presented in terms of the difference through time
and across sites, relative to the control site on the year of
implementation (\emph{i.e.} the difference-in-differences estimate or
effect size \(\lambda_t\) from Eqs. \ref{eqn:reg_bio} and
\ref{eqn:soc_reg}). We also provide an overview of the governance
settings of each community, and discuss how these might be related to
the effectiveness and performance of the reserves.

\hypertarget{biological-effects}{%
\subsection{Biological effects}\label{biological-effects}}

Indicators showed ambiguous responses through time for each reserve.
Figure \ref{fig:indicators}A shows positive effect sizes for lobster
densities in Isla Natividad and Punta Herrero during the first years,
but the effect is eroded through time. In the case of Maria Elena,
positive changes were observed in the third and fourth year. These
effects are in the order of 0.2 extra organisms \(\mathrm{m}^{-2}\) for
Isla Natividad and Punta Herrero, and 0.01 organisms \(\mathrm{m}^{-2}\)
for Maria Elena, but are not significantly different from zero
(\(p > 0.05\)). Likewise, no significant changes were detected in fish
biomass or invertebrate and fish densities (Fig.
\ref{fig:indicators}B-D), where effect sizes oscillated around zero
without clear trends. Figures and tables with time series of indicators
and model coefficients are presented in the supplementary materials
(Figures S1-S4, Tables S4-S6).

\hypertarget{socioeconomic-effects}{%
\subsection{Socioeconomic effects}\label{socioeconomic-effects}}

Lobster landings and revenue were only available for Isla Natividad and
Maria Elena (Fig \ref{fig:lobsters}). For all years before
implementation, the effect sizes are close to zero, indicating that the
control and treatment sites have similar pre-treatment trends,
suggesting that these are plausible controls. However, effect sizes do
not change after the implementation of the reserve. Interestingly, the
negative effect observed for Isla Natividad on year 5 correspond to the
2011 hypoxia events. The only positive change observed in lobster
landings is for Isla Natividad in 2014 (\(p < 0.1\)). The three years of
post-implementation data for Maria Elena do not show a significant
effect of the reserve. Isla Natividad shows higher revenues after the
implementation of the reserve, as compared to the control communities.
However, these changes are only significant for the third year
(\(p < 0.05\)). Full tables with model coefficients are presented in the
supplementary materials (Tables S4-S5).

\hypertarget{governance}{%
\subsection{Governance}\label{governance}}

Our analysis of the SES (Table \ref{table:ses}) shows that all analyzed
communities share similarities known to foster sustainable resource
management and increase reserve effectiveness. For example, fishers
operate within clearly outlined TURFs (RS2, GS6.1.4.3) that provide
exclusive access to resources and reserves. Along with their relatively
small groups (A1 - Number of relevant actors), Isolation (A3),
Operational rules (GS6.2), Social monitoring (GS9.1), and Graduated
sanctions (GS10.1), these fisheries have solid governance structures
that enable them to monitor their resources and enforce rules to ensure
sustainable management. In general, success of conservation initiatives
depends on the incentives of local communities to maintain a healthy
status of the resources upon which they depend \citep{jupiter_2017}. Due
to the clarity of access rights and isolation, the benefits of
conservation directly benefit the members of the fishing cooperatives,
which have favored the development of efficient community-based
enforcement systems. However, our SES analysis also highlights factors
that might hinder reserve performance or mask outcomes. While total
reserve size ranges from 0.2\% to 3.7\% of the TURF area, individual
reserves are often small (RS3); the largest reserve is only 4.37 km
\textsuperscript{2}, and the smallest one is 0.09 km
\textsuperscript{2}. Reserves are also relatively young (RS5).
Additionally, fishers harvest healthy stocks (RS4.1), and it's unlikely
that marine reserves will result in increased catches.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Our results indicate that these TURF-reserves have not increased lobster
densities. Additionally, no co-benefits were identified when using other
ecological indicators aside from the previously reported buffering
effect that reserves can have to environmental variability in Isla
Natividad \citep{micheli_2012-EU}. The socioeconomic indicators
pertaining landings and revenues showed little to no change after
reserve implementation. Lastly, the communities exhibit all the social
enabling conditions for effective reserve and resource management. Here
we discuss possible shortcomings in our analyses as well as possible
explanations for the observed patterns.

While many ecology studies have used BACI sampling designs and
respective analyses (\emph{e.g.} \citet{stewartoaten_1986}), few
conservation studies have done so to evaluate the effect of an
intervention (\emph{e.g.}
\citet{francinifilho_2008,lester_2009,moland_2013}) which has resulted
in a call for more robust analyses in conservation science
\citep{guidetti_2002,ferraro_2006}. Our approach to evaluate the
temporal and spatial changes provides a more robust measure of reserve
effectiveness. For example, we capture previously described patterns
like the rapid increase observed for lobster densities in Isla Natividad
on the sixth year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A), a year
after the hypoxia events described by \citet{micheli_2012-EU}, which
caused mass mortality of sedentary organisms such as abalone and sea
urchins, but not lobster and finfish. Yet, our empirical approach
assumes control sites are a plausible counterfactual for treated sites.
This implies that treated sites would have followed the same trend as
control sites, had the reserves not been implemented. Nonetheless,
temporal trends for each site don't show any significant increases
(Figures S1-S4), supporting our findings of lack of change in the
indicators used.

Our analyses of socioeconomic indicators has two limitations. First, we
only look at landings and revenues by landings for communities with and
without TURF-reserves. There are a number of other possible indicators
that could show a change due to the implementation of the reserve.
Notably, one often cited in the literature is additional benefits, such
as tourism \citep{viana_2017}. However, it is unlikely that the
evaluated communities will experience tourism benefits due to their
remoteness and the lack of proper infrastructure to sustain tourism. A
second limitation of our socioeconomic analysis is that we do not
observe effort data, which may mask the effect of the reserve. For
example, if catches remain relatively unchanged but fishing effort
decreased, that would imply a larger catch per unit effort and higher
profitability.

A first possible explanation for the lack of effectiveness may be the
young age of the reserves. Literature shows that age and enforcement are
important factors that influence reserve effectiveness
\citep{edgar_2014-UO,babcock_2010}. Isla Natividad has the oldest
reserves, and our SES analysis suggests that all communities have a
well-established community-based enforcement system. With these
characteristics, one would expect the reserves to be effective. Maria
Elena and Punta Herrero are relatively young reserves (\emph{i.e.}
\textless{} 6 years old) and effects may not yet be evident due to the
short duration of protection, relative to the life histories of the
protected species; community-based marine reserves in tropical
ecosystems may take six years or more to show a spillover effect
\citep{dasilva_2015-zX}.

Another key condition for effectiveness is reserve size
\citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be
attributed to poor ecological coherence in reserve design (\emph{sensu}
\citet{rees_2018}). Previous research has shown that reserves in Isla
Natividad yield fishery benefits for the abalone fishery
\citep{rossetto_2015-V0}. Abalone are less mobile than lobsters, and
perhaps the reserves provide enough protection to these sedentary
invertebrates, but not lobsters. Design principles developed by
\citet{green_2017} for marine reserves in the Caribbean state that
reserves ``should be more than twice the size of the home range of
adults and juveniles'', and suggest that reserves seeking to protect
spiny lobsters should have at least 14 km across. Furthermore, fishers
may favor implementation of reserves that pose low fishing costs due to
their small size or location. Our analysis of economic data supports
this hypothesis, as neither landings nor revenues showed the expected
short-term reductions associated to the first years of reserve
implementation \citep{ovando_2016-Wg}.

Even if reserves had appropriate sizes and were placed in optimal
locations, there are other plausible explanations for the observed
patterns. For instance, marine reserves are only likely to provide
fisheries benefits if initial population sizes are low and the fishery
is poorly managed \citep{hilborn_2004,hilborn_2006}. Both lobster
fisheries were certified by the Marine Stewardship Council and lobster
fisheries are managed via species-specific minimum catch sizes, seasonal
closures, protection of ``berried'' females, and escapement windows
where traps are allowed \citep{dof_website_1993}. It is uncertain
whether such a well-managed fishery will experience additional benefits
from marine reserves; reserves implemented in TURFs where fishing
pressure is already optimally managed will still show a trade-off
between fisheries and conservation objectives \citep{lester_2017}.
Furthermore, \citet{gelcich_2008} have shown that TURFs alone can have
greater biomass and richness than areas operating under open access.
This might reduce the difference between indicators from the TURF and
reserve sites, making it difficult to detect such a small change.
Further research should focus on evaluating sites in the reserve, TURF,
and open access areas or similar Fish Refuges established without the
presence of TURFs where the impact of the reserves might be larger.

Finally, extreme conditions, including prolonged hypoxia, heat waves,
and storms have affected both the Pacific and Caribbean regions, with
large negative impacts of coastal marine species and ecosystems
\citep{cavole_2016,hughes_2018,breitburg_2018}. The coastal ecosystems
where these reserves are located have been profoundly affected by these
events \citep{micheli_2012-EU,woodson_2018}. Effects of protection might
be eliminated by the mortalities associated with these extreme
conditions.

While the evaluated reserves have failed to provide fishery benefits up
to now, there are a number of additional ecological, fisheries, and
social benefits. Marine reserves provide protection to a wider range of
species and vulnerable habitat. Previous research focusing on these
specific sites has shown that they serve as an insurance mechanism
against uncertainty and errors in fisheries management, as well as mild
environmental shocks
\citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}.
Self-regulation of fishing effort can serve as a way to compensate for
future declines associated to environmental variation
\citep{finkbeiner_2018}. Furthermore, embarking in a marine conservation
project can bring the community together, which promotes social cohesion
and builds social capital as reported by \citet{fulton_2019}. Showing
commitment to marine conservation and sustainable fishing practices has
allowed fishers to have greater bargaining power and leverage over
fisheries management \citep{prezramrez_2012}. These additional benefits
might explain why communities show a positive perception about their
performance and continue to support their presence by re-implementing
the reserves \citep{ayer_2018}.

In terms of the fisheries regulation in Mexico, our work only evaluates
Fish Refuges established within TURFs. Future research should aim at
evaluating other Fish Refuges that have also been established as
bottom-up processes but without the presence of TURFs (\emph{e.g.}
\citet{dof_websiteC_2012}), others established through top-down
processes (\emph{i.e.} \citet{dof_websiteU_2018}), as well as the
relationship between governance and effectiveness across this gradient
of approaches.

Community-based TURF-reserves in small-scale fisheries may be helpful
conservation and fishery management tools when appropriately implemented
\citep{gelcich_2015}. We must promote bottom-up design and
implementation processes like the ones in the evaluated reserves, but
without setting design principles aside. Having full community support
surely represents an advantage, but it is important that community-based
TURF-reserves meet essential design principles such as size and
placement so as to maximize their effectiveness. For the particular case
of the ones we evaluate, the possibility of expanding reserves or
merging existing polygons into larger areas should be evaluated and
proposed to the communities.

\section*{Conflict of Interest Statement}

The authors declare that the research was conducted in the absence of
any commercial or financial relationships that could be construed as a
potential conflict of interest.

\section*{Author Contributions}

JC and AS conceived the idea. JC and EA analyzed data, discussed the
results, and wrote the first draft. FM, SF, AS, JT, and AHV discussed
the results and edited the manuscript. All authors provided valuable
contributions.

\section*{Funding}

JCVD received funding from UCMexus - CONACyT Doctoral Fellowship (CVU
669403) and the Latin American Fisheries Fellowship Program. AS, AHV, SF
and JT received funding from Marisla Foundation, Packard Foundation,
Walton Family Foundation, Summit Foundation, and Oak Foundation. FM was
supported by NSF-CNH and NSF BioOce (grants DEB-1212124 and 1736830).

\section*{Acknowledgments}

The authors wish to acknowledge Imelda Amador for contributions on the
governance data, as well as pre-processing biological data. This study
would have not been possible without the effort by members of the
fishing communities here mentioned, who participated in the
data-collection process.

\clearpage

\bibliographystyle{frontiersinSCNS_ENG_HUMS}
\bibliography{references}

\clearpage

\section*{Figure captions}

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{\label{fig:unnamed-chunk-7}\label{fig:map}Location of the three
coastal communities studied (A). Isla Natividad (B) is located off the
Baja California Peninsula, Maria Elena and Punta Herrero (C) are located
in the Yucatan Peninsula. Blue polygons represent the TURFs, and red
polygons the marine reserves.}
\end{figure}

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-8-1.pdf}
\caption{\label{fig:unnamed-chunk-8}\label{fig:indicators}Effect sizes for
marine reserves from Isla Natividad (IN; red circles), Maria Elena (ME;
blue triangles), and Punta Herrero (PH; green squares) for lobster
densities (\emph{Panulirus spp}; A), fish biomass (B), invertebrate
densities (C), and fish densities (D). Plots are ordered by survey type
(left column: invertebrates; right column: fish). Points are jittered
hotizontally to avoid overplotting. Points indicate the effect size and
error bars are heteroskedastic-robust standard errors. Years have been
centered to year of implementation.}
\end{figure}

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-9-1.pdf}
\caption{\label{fig:unnamed-chunk-9}\label{fig:lobsters}Effect sizes for
lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
circles) and Maria Elena (ME; blue triangles). Points are jittered
hotizontally to avoid overplotting. Points indicate the effect size and
error bars are heteroskedastic-robust standard errors. Years have been
centered to year of implementation.}
\end{figure}

\begin{table}[H]

\caption{\label{tab:unnamed-chunk-10}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
\centering
\begin{tabular}[t]{l|l}
\hline
Indicator & Units\\
\hline
\multicolumn{2}{l}{\textbf{Biological}}\\
\hline
\hspace{1em}Lobster density & org $\mathrm{m}^{-2}$\\
\hline
\hspace{1em}Invertebrate density & org $\mathrm{m}^{-2}$\\
\hline
\hspace{1em}Fish density & org $\mathrm{m}^{-2}$\\
\hline
\hspace{1em}Fish biomass & Kg $\mathrm{m}^{-2}$\\
\hline
\multicolumn{2}{l}{\textbf{Socioeconomic}}\\
\hline
\hspace{1em}Income from target species & M MXP\\
\hline
\hspace{1em}Landings from target species & Metric Tonnes\\
\hline
\end{tabular}
\end{table}

\begin{table}[H]

\caption{\label{tab:unnamed-chunk-11}\label{table:ses}Variables for the Social-Ecological System analysis (IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero). Alphanumeric codes follow \citet{basurto_2013-oq}; an asterisk (*) denotes variables incorporated based on \citet{difranco_2016-Xw} and \citet{edgar_2014-UO}. The presented narrative applies equally for all communities unless otherwise noted.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{6.5cm}|>{\raggedright\arraybackslash}p{12cm}}
\hline
Variable & Narrative\\
\hline
\multicolumn{2}{l}{\textbf{Resource System (RS)}}\\
\hline
\hspace{1em}RS2 - Clarity of system boundaries: Clarity of geographical boundaries of TURF and reserves & Individual TURF and reserve boundaries are explicitly outlined in official documents that include maps and coordinates. Reserve placement is decided by the community. Fishers use GPS units and landmarks.\\
\hline
\hspace{1em}RS3 - Size of resource system: TURF Area (Km$^2$) & IN = 889.5; ME = 353.1; PH = 299.7\\
\hline
\hspace{1em}RS3 - Size of resource system: Reserve area (Evaluated reserve area; Km$^2$) & IN = 2 (1.3); ME = 10.48(0.09); PH = 11.25 (4.37)\\
\hline
\hspace{1em}RS4.1 - Stock status: Status of the main fishery & Lobster stocks are well managed, and are (IN) or have been (ME, PH) MSC certified.\\
\hline
\hspace{1em}*RS5 - Age of reserves: Years since reserves were implemented & IN = 12; ME = 6; PH = 5\\
\hline
\multicolumn{2}{l}{\textbf{Resource Unit (RU)}}\\
\hline
\hspace{1em}RU5 - Number of units (catch diversity): Number of targeted species & Lobster is their main fishery of these three communities, but they also target finfish (2 spp each). Additionally, fishers from Isla Natividad target other sedentary benthic invertebrates (4 spp).\\
\hline
\multicolumn{2}{l}{\textbf{Actors (A)}}\\
\hline
\hspace{1em}A1 - Number of relevant actors: Number of fishers & IN = 98; ME = 80; PH = 21\\
\hline
\hspace{1em}*A3 - Isolation: Level of isolation of the fishing grounds & Their fishing grounds and reserves are highly isolated and away from dense urban centers. IN lies 545 Km south from Tijuana, and ME and PH 230 Km south from Cancun, where the nearest international airports are located.\\
\hline
\multicolumn{2}{l}{\textbf{Governance system (G)}}\\
\hline
\hspace{1em}GS6.1.4.3 - Territorial use communal rights : Presence of institutions that grant exclusive harvesting rights & Each community has exclusive access to harvest benthic resources, including lobster. These take the form of Territorial User Rights for Fisheries granted by the government to fishing cooperatives.\\
\hline
\hspace{1em}GS6.2 - Operational rules: Rules implemented by individuals atuhorized to partake on collective activities & Fishers have rules in addition to what the legislation mandates. These are: larger minimum catch sizes, lower quotas, and assigning fishers to specific fishing grounds within their TURF.\\
\hline
\hspace{1em}GS9.1 - Social monitoring: Monitoring of the activities performed by cooperative members and external fishers & Fishing cooperatives have a group (Consejo de vigilancia) that monitors and enforces formal and internal rules. They ensure fishers of their fishing cooperative adhere to the established rules, and that foreign vessels do not poach their TURF and reserves.\\
\hline
\hspace{1em}GS9.2 - Biophysical monitoring: Monitoring of biological resources, including targeted species & Fishers perform annual standardized underwater surveys in the reserves and fishing grounds. Recently, they have installed oceanographic sensors to monitor oceanographic variables.\\
\hline
GS10.1 - Graduated sanctions & Fishers have penalties for breaking collective-choice rules or fishing inside the reserves. These may range from scoldings and warnings to not being allowed to harvest a particular resource or being expelled from the cooperative.\\
\hline
\end{tabular}}
\end{table}

\end{document}