### 2019-02-28 docs/tab/S3_table.tex

\begin{table}[t]

\caption{\label{tab:}Mean ex-vessel prices (MXP / Kg) for lobster (Panulirus spp.) on each TURF (community) and group (Control = only TURF, Treated = TURF and reserve).}
\centering
\begin{tabular}{lllrr}
\toprule
Community & Group & TURF & Before & After\\
\midrule
Yucatan Peninsula & Control & Langosteros Del Caribe & 165.58 & 176.72\\
Yucatan Peninsula & Control & Vigia Chico & 177.66 & 175.31\\
Yucatan Peninsula & Treated & Cozumel & 178.21 & 170.49\\
Baja California & Control & Bahia Tortugas & 155.60 & 122.46\\
Baja California & Control & La Purisima & 67.54 & 157.88\\
\addlinespace
Baja California & Control & Pesc Nacionales De Abulon & 67.16 & 88.79\\
Baja California & Treated & Buzos Y Pescadores De La Baja California & 89.35 & 190.07\\
\bottomrule
\end{tabular}
\end{table}